---
title: 'Applied Data Science Using R - INF 2167H - Assignment #1'
author: "Faria Khandaker"
output:
  pdf_document:
    latex_engine: lualatex
  html_document:
    df_print: paged
font: 12pt
Due date: October 14, 2020
---
Use RStudio for this assignment. 
Edit the file `assignment01_Fall2020.Rmd` and insert your R code wherever you see the string "INSERT YOUR ANSWER HERE"

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. 
To be able to read the questions well, knit the document into word format first.

## Sample Question and Solution

Use `seq()` to create the vector $(1,2,3,...,10)$.

```{r}
seq(1,10)
```
To properly read some questions (ex: question 2), knit the document to word or pdf before you start solving the assignment

## Question 1
In the following exercises of question 1, use only seq, rep, or c

a) (1.5 points) Create the vector (5, 9, 13, ..., 41). Note that each term in this sequence is of the form 
1 + 4n where n = 1, ..., 10.
```{r}

n=1:10
a<-1+4*n
c(a)

```

b) (1.5 points) Create the vector (2, 3, 4, ..., 10, 9, 8, ..., 2).
```{r}
m=c(seq(2,10),sort(seq(9,2), decreasing = T))
m

```


c) (1.5 points) Create the vector (1,2,3,...,1,2,3) in which the sequence (1,2,3) is repeated 5 times.
```{r}
rep(c(1,2,3),5)
```


d) (1.5 points) Create the vector (1,1,...,1,2,2,...,2,3,3,...,3) where each number is repeated 7 times.
```{r}
a=1
b=2
c=3
num=c(rep(1,7),rep(2,7),rep(3,7))
num

```

e) (1.5 points) Create the vector (10,20,20,30,30,30,...,100,...,100) where 10n is repeated n times.
```{r}

y=1:10
z<-c(rep(10*y,y))

z

```


## Question 2

a) (1.5 points) Compute: $$\sum_{n=1}^{100} n$$
```{r}
sum(1:100)

```


b) (1.5 points) Compute: $$\sum_{n=1}^{100} n^2$$
```{r}
n=seq(100)
sum(n^2)

```


c) (1.5 points) Compute: $$\sum_{n=10}^{20} \left(\frac{2^{n}}{n} + \frac{3^{n}}{n^{3}}\right)$$
```{r}
d=seq(10,20)
sum(((2^d)/d)+((3^d)/d^3))

```


d) (1.5 points) Compute: $$\sum_{n=0}^{10} \frac{1}{n!}$$ 
Hint: Use `factorial(n)` to compute $n!$
```{r}
e=seq(10)
sum(1/factorial(e))

```


e) (1.5 points) Compute: $$\sum_{n=1}^{20} \left(2n + \frac{1}{n^2}\right)$$
```{r}
f=seq(20)
sum((2*f)+(1/(f^2)))

```


## Question 3

a) (1.5 point) Create an empty list `mylist`.
```{r}
mylist<-list()
mylist

```


b) (1.5 points) Add a component named `aa` whose value is 42.
```{r}
mylist<-list(mylist,"aa"=42)



```


c) (2.5 points) Add a component named `bb` whose value is the numeric vector $(1,2,\ldots,10)$.
```{r}
mylist<-list(mylist,"bb"=c(1:10))


```

d) (2 points) Add a component named `cc` whose value is the character vector ("Hello", "INF 2167").
```{r}
mylist<-list(mylist, "cc"= c("Hello", "INF 2167"))

```


e) (2 points) Add a component named `dd` whose value is a 4x3 matrix whose elements are $(1,2,\ldots,12)$ in column-major order.
```{r}
mat <- matrix(1:12, 4, 3)
mylist<-list(mylist, "dd"= mat)

```


f) (0.5 point) Display `mylist` on the screen.
```{r}
mylist
```


## Question 4

If you have not already done so, install the ISwR package on
your computer using the command `install.packages("ISwR")`.
Note that install.packages causes errors while knitting so make sure after you install the package to hash the command  (i.e. #install.packages("ISWR"))

Loading the `ISwR` package into the current session.
```{r}
#install.packages("ISwR")
library(ISwR)

```

a) (1 points) Display the head of the `thuesen` data frame.
```{r}
dataset<-data("thuesen")
head(thuesen)
```


b) (4 points) Compute the mean of each variable using `sapply()`, removing the missing values.
```{r}

nomissing <- na.omit(thuesen)
nomissing$blood.glucose<-as.numeric(nomissing$blood.glucose)
nomissing$short.velocity<-as.numeric(nomissing$short.velocity)
sapply(nomissing,function(x) mean(x))

```


c) (3 points) Create a numeric vectors `n1`, `n2`, and `n3` whose elements are the integers from 1 to 20, their squares, and their cubes.
```{r}
n1=0
n2=0
n3=0
d1=seq(20)
n1=c(d1)
n2=c(d1^2)
n3=c(d1^3)
```


d) (2 points) Create a new data frame `nn` from the above three vectors.
```{r}
nn=data.frame(n1,n2,n3)
```


e) (1 points) Display the tail of `nn`.
```{r}
tail(nn)
```


f) (4 points) Compute the sum of each variable in `nn` using `sapply`.
```{r}
sapply(nn, function(x) sum(x))

```

## Question 5


a) (3 points) Create a 4x4 empty matrix, i.e. all elements equal to NA, display mat1. 

```{r}
mat1<- matrix(NA, 4,4)
mat1

```
b) (7 points) fill the middle 4 elements with the values 'This' 'is' 'the' 'middle' and display mat1.
```{r}

mat1[2,2]<-"This"
mat1[3,2]<-"is"
mat1[2,3]<-"the"
mat1[3,3]<-"middle"
mat1

```
c) (10 points) This is the code to write the "This is the middle" row-wise
```{r}

mat2<- matrix(NA, 4,4)
mat2[2,2]<-"This"
mat2[2,3]<-"is"
mat2[3,2]<-"the"
mat2[3,3]<-"middle"
mat2
```

## Question 6
Use the tidyverse library
a) (2 points) Import the dataset WineQuality (4898 rows x12 columns) available in the following website
http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv, note that the type of separator in this dataset = ;
```{r message=FALSE, warning=FALSE}
library(tidyverse)
winequality<-read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", sep = ";")

```
b) (2 points) Add a column named total.acidity which is the sum of fixed.acidity and volatile.acidity
Display the head of the new column
```{r}
winequality<-mutate(winequality, total.acidity= fixed.acidity+volatile.acidity)
head(winequality$total.acidity)

```
c) (4 points) Create a dataframe named bestWine to only include wine of best quality 
How many cases have wine of best quality?
A wine of best quality is wine that has highest value in quality
```{r}
#max(winequality$quality)
bestWine<- filter(winequality,winequality$quality >= 9)
nrow(bestWine)
```
d) (4 points) Display the head of bestWine showing only the pH, alcohol, and quality
```{r}
head(select(bestWine, pH,alcohol,quality))

```
e) (5 points) Sort the WineQuality dataset in descending order of alcohol concentration 
Display only the tail of the alcohol, quality and density 
```{r}
sortedWine<-arrange(winequality, desc(winequality$alcohol))
tail(select(sortedWine, alcohol, quality, density))

```
f) (5 points) Add a variable to the data frame that takes value 1 if the food has higher citric acid than average, 0 otherwise.Call this variable HighAcid. Do the same for High chlorides, High sugar, and High sulphates. How many cases have both high acid and high chlorides? 
```{r}
#calculating the means of sugar, citric acid and sulphates
avacid<-mean(winequality$citric.acid)
avchlor<-mean(winequality$chlorides)
avsug<-mean(winequality$residual.sugar)
avsul<-mean(winequality$sulphates)

#for high citric acid
winequality<-mutate(winequality, high.citric.acid= ifelse(citric.acid>avacid,1,0))

#for high chloride
winequality<-mutate(winequality, high.chloride= ifelse(chlorides>avchlor,1,0))

#for high sugar
winequality<-mutate(winequality, high.sugar= ifelse(residual.sugar>avsug,1,0))

#for high sulphates
winequality<-mutate(winequality, high.sulphate= ifelse(sulphates>avsul,1,0))

#high chlorides and high acid

sum(winequality$high.chloride & winequality$high.citric.acid == 1)
```
g) (8 points) Create a function called Wine.check to detect bad quality wine. Use the flowchart attached to the assignment as a basis for this function. 
Hint: Use nested if statement inside the function
```{r}
Wine.check <- function(m){
  ifelse(winequality$high.citric.acid==0, "Pass", ifelse(winequality$high.chloride==0, "Pass",
                                                         ifelse(winequality$high.sugar==0, "Pass", "Fail")))
}

```

h) (7 points) Create a new variable called WineCheck using the output of the function. This variable will have values of either "pass" or "Fail". A "Fail" will occur when all acid, chlorides and sugar are high (=1), display the head of WineCheck
```{r}
Winecheck<-Wine.check(winequality)
head(Winecheck)

```
i) (3 points) How many wines in the WineQuality data frame fail the WineCheck? (8 points)
```{r}
winequality<-mutate(winequality, winecheck=Wine.check(winequality1))
sum(winequality$winecheck=="Fail")

```

END of Assignment #1. Good Luck!

