---
title: "Faria Individual Submission"
author: "Faria"
date: "12/9/2020"
output: html_document
---
```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(rcompanion)
library(MASS)# stepwise regression
library(randomForest) 
library(leaps) # all subsets regression
library(corrplot)
library(caret)
library(FNN)
library(mlbench)
library(lattice)
library(class)
library(gmodels)
library(mice)
```
# Question 2 part 1 Data cleaning and Models
```{r, Weeklypharmasalesdatacleaning}
#run this
# Pharma sales dataset
#do linear regression to check of correlation for between and within years
weeklypharma<-read.csv("salesweekly.csv")
colnames(weeklypharma)
colnames(weeklypharma)<-c("Date","Med4RheumArth","Med4OstArth","Aspirin","Ibuprofen",
                          "Med4Tension","Med4Sleep","Med4Asthma","Med4Allergy")
str(weeklypharma)
#convert to date time object
weeklypharma$Date<-mdy(weeklypharma$Date)

#Obama year sales
pharma201415<-weeklypharma
pharma201415 <- 
  pharma201415 %>% 
  filter(between(Date,as.Date("2014-01-01"), as.Date("2015-12-31"))) 

#Trump year sales
pharma201718<-weeklypharma
pharma201718 <- 
  pharma201718 %>% 
  filter(between(Date,as.Date("2017-01-01"), as.Date("2018-12-31")))

#election year sales
pharma2016<-weeklypharma
pharma2016 <- 
  pharma2016 %>% 
  filter(between(Date,as.Date("2016-01-01"), as.Date("2016-12-31")))
```

```{r obama data exploration and transformation for linear reg}
# first, load cleaned datasets from FinalProjectDataCleaning
obamapharma<-pharma201415[,-1]
correlationMatrix <- cor(obamapharma, method = "pearson")
correlationMatrix
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(correlationMatrix, method="color", col=col(200),type="upper", order="hclust")
# MEdicine are not highly correlated (all meds were 0.4 or under) with each other therefore each no med was dropped. 
colnames(pharma201415)
str(pharma201415)
library(e1071)                    # load e1071 
skewness(pharma201415$Med4Sleep)
#plotNormalHistogram(pharma201415$Meds4Allergy)
#no missing values in the phamadata
#med4rheumarth skewness is 0.1803
#med4ostarth skew is -0.025
#med4sleep skew is 0.897, rskew
#med4asthma skew is 0.97, r skew
#med4allergy skew is 0.9579, rskew
#med4tension is 0.793, rskew
#ibuprofen is 0.79731, rskew
#aspirin is 0.5888, right skewed
#6/8 variables are right skewed, log transformation is done
#("Date","Med4RheumArth","Med4OstArth"," Aspirin","Ibuprofen"," Med4Tension","Med4Sleep","Meds4Asthma","Meds4Allergy")

#log transformations required, add small coefficient (0.5) to prevent log of 0
#do them individually so that the dataframe isnt spliced
dolog<-function(x){
  log(x+1)
}
t201415<-pharma201415
t201415[2:9]<-lapply(t201415[2:9],dolog)#logtransformed

```


# Stepwise Feature Selection
Since none of them were correlated, hard to know which features are best predictors of independent variables
```{r feature selection for rheumatoid arthritis meds}
fullobamarh <- lm(Med4RheumArth~., data= t201415)  ## . means all the IVs
nullobamarh <- lm(Med4RheumArth~1,data=t201415)
stepobamarh <- stepAIC(nullobamarh, scope=list(lower=nullobamarh, upper=fullobamarh), direction= "both", trace=FALSE)
summary(stepobamarh) 
```

```{r feature selection for osteo arthritis meds}
fullobamaost <- lm(Med4OstArth~., data= t201415)  ## . means all the IVs
nullobamaost <- lm(Med4OstArth~1,data=t201415)
stepobamaost <- stepAIC(nullobamaost, scope=list(lower=nullobamaost, upper=fullobamaost), direction= "both", trace=FALSE)
summary(stepobamaost) 

```

```{r feature selection for aspirin}
fullobamaasp <- lm(Aspirin~., data= t201415)  ## . means all the IVs
nullobamaasp <- lm(Aspirin~1,data=t201415)
stepobamaasp <- stepAIC(nullobamaasp, scope=list(lower=nullobamaasp, upper=fullobamaasp), direction= "both", trace=FALSE)
summary(stepobamaasp) 
```

```{r feature selection for ibuprofen}
fullobamaib <- lm(Ibuprofen~., data= t201415)  ## . means all the IVs
nullobamaib <- lm(Ibuprofen~1,data=t201415)
stepobamaib <- stepAIC(nullobamaib, scope=list(lower=nullobamaib, upper=fullobamaib), direction= "both", trace=FALSE)
summary(stepobamaib) 

```

```{r feature selection for tension medicine}
fullobamaten <- lm(Med4Tension~., data= t201415)  ## . means all the IVs
nullobamaten <- lm(Med4Tension~1,data=t201415)
stepobamaten <- stepAIC(nullobamaten, scope=list(lower=nullobamaten, upper=fullobamaten), direction= "both", trace=FALSE)
summary(stepobamaten) 
```

```{r feature selection for sleep medicine}
fullobamaslp <- lm(Med4Sleep~., data= t201415)  ## . means all the IVs
nullobamaslp <- lm(Med4Sleep~1,data=t201415)
stepobamaslp <- stepAIC(nullobamaslp, scope=list(lower=nullobamaslp, upper=fullobamaslp), direction= "both", trace=FALSE)
summary(stepobamaslp) 
```

```{r feature selection for asthma medicine}
fullobamaast <- lm(Med4Asthma~., data= t201415)  ## . means all the IVs
nullobamaast <- lm(Med4Asthma~1,data=t201415)
stepobamaast <- stepAIC(nullobamaast, scope=list(lower=nullobamaast, upper=fullobamaast), direction= "both", trace=FALSE)
summary(stepobamaast) 
```

```{r feature selection for allergy medicine}
fullobamaall <- lm(Med4Allergy~., data= t201415)  ## . means all the IVs
nullobamaall <- lm(Med4Allergy~1,data=t201415)
stepobamaall <- stepAIC(nullobamaall, scope=list(lower=nullobamaall, upper=fullobamaall), direction= "both", trace=FALSE)
summary(stepobamaall) 
```

# Trump Data
```{r trump data exploration and transformation for linear reg}
trumppharma<-pharma201718[,-1]
correlationMatrix <- cor(trumppharma, method = "pearson")
correlationMatrix
#the correlations are not strong but are higher than obama period
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(correlationMatrix, method="color", col=col(200),type="upper", order="hclust")
# MEdicine are not highly correlated (all meds were 0.4 or under) with each other therefore each no med was dropped. 
colnames(pharma201718)
str(pharma201718)
library(e1071)                    # load e1071 
#skewness(pharma201718$Aspirin)
#plotNormalHistogram(pharma201415$Aspirin)

#no missing values in the phamadata
#med4rheumarth skewness is 0.537, range (3.012-4.04)
#med4ostarth skew is 0.39, range(2.763674 3.798137)
#med4sleep skew is 1.4, rskew range(0.000000 2.89037)
#med4asthma skew is 0.699, r skew range(1.609438 4.762174)
#med4allergy skew is 1.15, rskew range(1.609438 4.189655)
#med4tension is 1,17, rskew range(2.944439 4.795791)
#ibuprofen is 1.10, rskew range(4.541698 6.306091)
#aspirin is 1.29 , right skewed range(2.411439 4.112921)

#7/8 variables are right skewed, log transformation is done
#("Date","Med4RheumArth","Med4OstArth"," Aspirin","Ibuprofen"," Med4Tension","Med4Sleep","Meds4Asthma","Meds4Allergy")

#log transformations required, add small coefficient (0.5) to prevent log of 0
#do them individually so that the dataframe isnt spliced
dolog<-function(x){
  log(x+1)
}
t201718<-pharma201718
t201718[2:9]<-lapply(t201718[2:9],dolog)#logtransformed
range(t201718$Aspirin)

```

# Trump Stepwise Feature Selection
Since none of them were correlated, hard to know which features are best predictors of independent variables
```{r feature selection for rheumatoid arthritis meds}
fulltrumprh <- lm(Med4RheumArth~., data= t201718)  ## . means all the IVs
nulltrumprh <- lm(Med4RheumArth~1,data=t201718)
steptrumprh <- stepAIC(nulltrumprh, scope=list(lower=nulltrumprh, upper=fulltrumprh), direction= "both", trace=FALSE)
summary(steptrumprh) 
```

```{r feature selection for osteo arthritis meds}
fulltrumpost <- lm(Med4OstArth~., data= t201718)  ## . means all the IVs
nulltrumpost <- lm(Med4OstArth~1,data=t201718)
steptrumpost <- stepAIC(nulltrumpost, scope=list(lower=nulltrumpost, upper=fulltrumpost), direction= "both", trace=FALSE)
summary(steptrumpost) 

```

```{r feature selection for aspirin}
fulltrumpasp <- lm(Aspirin~., data= t201718)  ## . means all the IVs
nulltrumpasp <- lm(Aspirin~1,data=t201718)
steptrumpasp <- stepAIC(nulltrumpasp, scope=list(lower=nulltrumpasp, upper=fulltrumpasp), direction= "both", trace=FALSE)
summary(steptrumpasp) 
```

```{r feature selection for ibuprofen}
fulltrumpib <- lm(Ibuprofen~., data= t201718)  ## . means all the IVs
nulltrumpib <- lm(Ibuprofen~1,data=t201718)
steptrumpib <- stepAIC(nulltrumpib, scope=list(lower=nulltrumpib, upper=fulltrumpib), direction= "both", trace=FALSE)
summary(steptrumpib) 

```

```{r feature selection for tension medicine}
fulltrumpten <- lm(Med4Tension~., data= t201718)  ## . means all the IVs
nulltrumpten <- lm(Med4Tension~1,data=t201718)
steptrumpten <- stepAIC(nullobamaten, scope=list(lower=nulltrumpten, upper=fulltrumpten), direction= "both", trace=FALSE)
summary(steptrumpten) 
```

```{r feature selection for sleep medicine}
fulltrumpslp <- lm(Med4Sleep~., data= t201718)  ## . means all the IVs
nulltrumpslp <- lm(Med4Sleep~1,data=t201718)
steptrumpslp <- stepAIC(nulltrumpslp, scope=list(lower=nulltrumpslp, upper=fulltrumpslp), direction= "both", trace=FALSE)
summary(steptrumpslp) 
```

```{r feature selection for asthma medicine}
fulltrumpast <- lm(Med4Asthma~., data= t201718)  ## . means all the IVs
nulltrumpast <- lm(Med4Asthma~1,data=t201718)
steptrumpast <- stepAIC(nulltrumpast, scope=list(lower=nulltrumpast, upper=fulltrumpast), direction= "both", trace=FALSE)
summary(steptrumpast) 
```

```{r feature selection for allergy medicine}
fulltrumpall <- lm(Med4Allergy~., data= t201718)  ## . means all the IVs
nulltrumpall <- lm(Med4Allergy~1,data=t201718)
steptrumpall <- stepAIC(nulltrumpall, scope=list(lower=nulltrumpall, upper=fulltrumpall), direction= "both", trace=FALSE)
summary(steptrumpall) 
```


NOTES: stepwise regression was done for weekly on the time periods 201415 and 201718. for the 201415 period, 5/8 meds had date as significant but only 1/8 meds had date as signifcant. 


# Training and testing 2014-2015 models to make sure model gives back good metrics before testing on other data

```{r traintest split }
#linear reg train test split
set.seed(111)
rn_train <- sample(nrow(t201415), floor(nrow(t201415)*0.7))
obamatrain70<-t201415[rn_train,]
obamatest<-t201415[-rn_train,]


```

Using Linear Regression to train 70% 201415 data and test 30% on 20142015
rheumArth, Aspirin,Ibuprofen,Sleep Med and Allergy Med had date be a significant feature so only those 5 meds were used for training and test.

```{r linear model for rheum arthritis}
ObamaRheumModel<-lm(Med4RheumArth~Date, data=obamatrain70) 
ObamaRheumPrediction <- predict(ObamaRheumModel, newdata =obamatest) 
ObamaRheumErrors <- ObamaRheumPrediction - obamatest$Med4RheumArth
hist(ObamaRheumErrors)
ObamaRheumRmse <- sqrt(mean((obamatest$Med4RheumArth - ObamaRheumPrediction)^2))
ObamaRheumRel_change <- abs(ObamaRheumErrors) / obamatest$Med4RheumArth
ObamaRheumPred10 <- table(ObamaRheumRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaRheumPred25 <- table(ObamaRheumRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Rheumatoid Arthritis Meds")
paste("RMSE:", ObamaRheumRmse)
paste("PRED(10):", round(ObamaRheumPred10,2))
paste("PRED(25):", round(ObamaRheumPred25,2))
paste("Summary of Prediction")
summary(ObamaRheumPrediction)


#88%of values are within 10% of the actual values, 100% of values are within 25% of actualvalues and there is a very low RMSE(0.178) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Aspirin}
ObamaAspModel<-lm(Aspirin~Date, data=obamatrain70) 
ObamaAspPrediction <- predict(ObamaAspModel, newdata =obamatest) 
ObamaAspErrors <- ObamaAspPrediction - obamatest$Aspirin
hist(ObamaAspErrors)
ObamaAspRmse <- sqrt(mean((obamatest$Aspirin - ObamaAspPrediction)^2))
ObamaAspRel_change <- abs(ObamaAspErrors) / obamatest$Aspirin

ObamaAspPred10 <- table(ObamaAspRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaAspPred25 <- table(ObamaAspRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Aspirin")
paste("RMSE:", ObamaAspRmse)
paste("PRED(10):", round(ObamaAspPred10,2))
paste("PRED(25):", round(ObamaAspPred25,2))
paste("Summary of Prediction")
summary(ObamaAspPrediction)

#within 201415
#88% of values are within 10% of the actual values and 100% of values are within 25% of actual values there is a very low RMSE(0.243) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Ibuprofen}

ObamaIbModel<-lm(Ibuprofen~Date, data=obamatrain70) 
ObamaIbPrediction <- predict(ObamaIbModel, newdata =obamatest) 
ObamaIbErrors <- ObamaIbPrediction - obamatest$Ibuprofen
hist(ObamaIbErrors)
ObamaIbRmse <- sqrt(mean((obamatest$Ibuprofen - ObamaIbPrediction)^2))
ObamaIbRel_change <- abs(ObamaIbErrors) / obamatest$Ibuprofen
ObamaIbPred10 <- table(ObamaIbRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaIbPred25 <- table(ObamaIbRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Ibuprofen")
paste("RMSE:", ObamaIbRmse)
paste("PRED(10):", round(ObamaIbPred10,2))
paste("PRED(25):", round(ObamaIbPred25,2))
paste("Summary of Prediction")
summary(ObamaIbPrediction)
#within201415

#88%of values are within 10% of the actual values and 100% of values fall within 25% of actual values (match 100%) and there is a very low RMSE(0.283) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for sleep med}
ObamaSlpModel<-lm(Med4Sleep~Date, data=obamatrain70) 
ObamaSlpPrediction <- predict(ObamaSlpModel, newdata =obamatest) 
ObamaSlpErrors <- ObamaSlpPrediction - obamatest$Med4Sleep
hist(ObamaSlpErrors)
ObamaSlpRmse <- sqrt(mean((obamatest$Med4Sleep - ObamaSlpPrediction)^2))
ObamaSlpRel_change <- abs(ObamaSlpErrors) / obamatest$Med4Sleep
ObamaSlpPred10 <- table(ObamaSlpRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaSlpPred25 <- table(ObamaSlpRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Sleep Meds")
paste("RMSE:", ObamaSlpRmse)
paste("PRED(10):", round(ObamaSlpPred10,2))
paste("PRED(25):", round(ObamaSlpPred25,2))
paste("Summary of Prediction")
summary(ObamaSlpPrediction)

#22%of values are within 10% of the actual values, 66% fall within 25% of the actual values and there is a very moderate RMSE(0.65) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for allergymed}

ObamaAllModel<-lm(Med4Allergy~Date, data=obamatrain70) 
ObamaAllPrediction <- predict(ObamaAllModel, newdata =obamatest) 
ObamaAllErrors <- ObamaAllPrediction - obamatest$Med4Allergy
hist(ObamaAllErrors)
ObamaAllRmse <- sqrt(mean((obamatest$Med4Allergy - ObamaAllPrediction)^2))
ObamaAllRel_change <- abs(ObamaAllErrors) / obamatest$Med4Allergy
ObamaAllPred10 <- table(ObamaAllRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaAllPred25 <- table(ObamaAllRel_change<0.25)["TRUE"] / nrow(obamatest)
paste("2014-2015 model for Allergy Meds")
paste("RMSE:", ObamaAllRmse)
paste("PRED(10):", round(ObamaAllPred10,2))
paste("PRED(25):", round(ObamaAllPred25,2))
paste("Summary of Prediction")
summary(ObamaAllPrediction)
#41%of values are within 10% of the actual values, 81.25% are within 25% of actual values and there is a very low RMSE(0.54) which may indicate that the sales of medicines follows a certain pattern.
```

# Obama train, trump test

Using Linear Regression to train 201415 data and test on 201718.
rheumArth, Aspirin,Ibuprofen,Sleep Med and Allergy Med had date be a significant feature so only those 5 meds were used for training and test.

```{r traintest split }
#linear reg train test split
#rn_train <- sample(nrow(t201415), floor(nrow(t201415)*0.7))
#obamatrain<-t20142015[rn_train,]
#obamatest<-t20142015[-rn_train,]

obamatrain<-t201415
trumptest<-t201718[-1,]#to keep the number of observations the same
# to reset index rownames(trumptest)<-NULL

```

```{r linear model for rheum arthritis}
OTRheumModel<-lm(Med4RheumArth~Date, data=obamatrain) 
OTRheumPrediction <- predict(OTRheumModel, newdata =trumptest) 
OTRheumErrors <- OTRheumPrediction - trumptest$Med4RheumArth
hist(OTRheumErrors)
OTRheumRmse <- sqrt(mean((trumptest$Med4RheumArth - OTRheumPrediction)^2))
OTRheumRel_change <- abs(OTRheumErrors) / trumptest$Med4RheumArth
OTRheumPred10 <- table(OTRheumRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTRheumPred25 <- table(OTRheumRel_change<0.25)["TRUE"] / nrow(trumptest)
paste("TrumpTest: Rheumatoid Athritis Med")
paste("RMSE:", OTRheumRmse)
paste("PRED(10):", round(OTRheumPred10,2))
paste("PRED(25):", round(OTRheumPred25,2))
paste("Summary of Prediction")
summary(OTRheumPrediction)
plot(OTRheumErrors)
plot(fitted(OTRheumModel), resid(OTRheumModel))
#91%of values are within 10% of the actual values, 100% of values are within 25% of actualvalues and there is a very low RMSE(0.178) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Aspirin}
OTAspModel<-lm(Aspirin~Date, data=obamatrain) 
OTAspPrediction <- predict(OTAspModel, newdata=trumptest) 
OTAspErrors <- OTAspPrediction - trumptest$Aspirin
hist(OTAspErrors)
OTAspRmse <- sqrt(mean((trumptest$Aspirin - OTAspPrediction)^2))
OTAspRel_change <- abs(OTAspErrors) / trumptest$Aspirin

OTAspPred10 <- table(OTAspRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTAspPred25 <- table(OTAspRel_change<0.25)["TRUE"] / nrow(trumptest)
paste("TrumpTest: Aspirin")
paste("RMSE:", OTAspRmse)
paste("PRED(10):", round(OTAspPred10,2))
paste("PRED(25):", round(OTAspPred25,2))
paste("Summary of Prediction")
summary(OTAspPrediction)

#60% of values are within 10% of the actual values and 91% of values are within 25% of actual values there is a very low RMSE(0.243) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Ibuprofen}

OTIbModel<-lm(Ibuprofen~Date, data=obamatrain) 
OTIbPrediction <- predict(OTIbModel, newdata =trumptest) 
OTIbErrors <- OTIbPrediction - trumptest$Ibuprofen
hist(OTIbErrors)
OTIbRmse <- sqrt(mean((trumptest$Ibuprofen - OTIbPrediction)^2))
OTIbRel_change <- abs(OTIbErrors) / trumptest$Ibuprofen
OTIbPred10 <- table(OTIbRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTIbPred25 <- table(OTIbRel_change<0.25)["TRUE"] / nrow(trumptest)

paste("TrumpTest: Ibuprofen")
paste("RMSE:", OTIbRmse)
paste("PRED(10):", round(OTIbPred10,2))
paste("PRED(25):", round(OTIbPred25,2))
paste("Summary of Prediction")
summary(OTIbPrediction)


# 26% of predicted values fall within 25% of actual values . 91%of values are within 25% of the actual values and there is a very low RMSE(0.283) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for sleep med}

OTSlpModel<-lm(Med4Sleep~Date, data=obamatrain) 
OTSlpPrediction <- predict(OTSlpModel, newdata=trumptest) 
OTSlpErrors <- OTSlpPrediction - trumptest$Med4Sleep
hist(OTSlpErrors)
OTSlpRmse <- sqrt(mean((trumptest$Med4Sleep - OTSlpPrediction)^2))
OTSlpRel_change <- abs(OTSlpErrors) / trumptest$Med4Sleep
OTSlpPred10 <- table(OTSlpRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTSlpPred25 <- table(OTSlpRel_change<0.25)["TRUE"] / nrow(trumptest)

paste("TrumpTest: Sleep Medicine")
paste("RMSE:", OTSlpRmse)
paste("PRED(10):", round(OTSlpPred10,2))
paste("PRED(25):", round(OTSlpPred25,2))
paste("Summary of Prediction")
summary(OTSlpPrediction)

#None of the values are within 10% of the actual values, 2% fall within 25% of the actual values and there is an RMSE 1.25 which is very high which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for allergymed}

OTAllModel<-lm(Med4Allergy~Date, data=obamatrain) 
OTAllPrediction <- predict(OTAllModel, newdata =trumptest) 
OTAllErrors <- OTAllPrediction - trumptest$Med4Allergy
hist(OTAllErrors)
OTAllRmse <- sqrt(mean((trumptest$Med4Allergy - OTAllPrediction)^2))
OTAllRel_change <- abs(OTAllErrors) / trumptest$Med4Allergy
OTAllPred10 <- table(OTAllRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTAllPred25 <- table(OTAllRel_change<0.25)["TRUE"] / nrow(trumptest)

paste("TrumpTest: Allergy Medicine")
paste("RMSE:", OTAllRmse)
paste("PRED(10):", round(OTAllPred10,2))
paste("PRED(25):", round(OTAllPred25,2))
paste("Summary of Predictions")
summary(OTAllPrediction)

#41%of values are within 10% of the actual values, 81.25% are within 25% of actual values and there is a very low RMSE(0.54) which may indicate that the sales of medicines follows a certain pattern.
```


# Question 2 part 2 Data cleaning and Models
```{r stocktickers found in 200+dataset}
#run this
nyseticker<-read.csv("nyseticker.csv")


nasdaqticker<-read.csv("nasdaqticker.csv")


#change names of pricevar column so it is same for all dataframes
st2002014<-read.csv("2014_Financial_Data.csv")
colnames(st2002014)[224]<-"NxtYearPriceVar"

st2002015<-read.csv("2015_Financial_Data.csv")
colnames(st2002015)[224]<-"NxtYearPriceVar"

st2002016<-read.csv("2016_Financial_Data.csv")
colnames(st2002016)[224]<-"NxtYearPriceVar"

st2002017<-read.csv("2017_Financial_Data.csv")
colnames(st2002017)[224]<-"NxtYearPriceVar"

st2002018<-read.csv("2018_Financial_Data.csv")
colnames(st2002018)[224]<-"NxtYearPriceVar"
unique(st2002014$Sector)


obamastocks<-rbind(st2002014,st2002015)
obamahealthcare<-obamastocks[obamastocks$Sector=="Healthcare", ]

trumpstocks<-rbind(st2002017,st2002018)
trumphealthcare<-trumpstocks[trumpstocks$Sector=="Healthcare", ]

trumphealthcare$nyselabel<-ifelse(trumphealthcare$X%in%nyseticker$Ticker,1,0)
trumphealthcare$naslabel<-ifelse(trumphealthcare$X%in%nasdaqticker$Ticker,1,0)
obamahealthcare$nyselabel<-ifelse(obamahealthcare$X%in%nyseticker$Ticker,1,0)
obamahealthcare$naslabel<-ifelse(obamahealthcare$X%in%nasdaqticker$Ticker,1,0)

```
# only obama model

```{r}
#str(obamahealthcare)
#get rid of columns that will give feature selection trouble
obamahealth<-obamahealthcare[,-c(93,112,114,223)]
```

```{r impute missing values}
for(i in 1:ncol(obamahealth)) {
  obamahealth[ , i][is.na(obamahealth[ , i])] <- mean(obamahealth[ , i], na.rm = TRUE)
}
#label the dataset in the data cleaning rmd
#now only pharmaceutical on the stock exchanges are included
notpharma<-c(which(obamahealth$naslabel==0 & obamahealth$nyselabel==0))
obamahealth<-obamahealth[-notpharma,]
sum(is.na(obamahealth))
```
```{r correlation and normalization}
#log everything to normalize dataset, add high constant 
normfunc<-function(x){
  (x-min(x))/(max(x)-min(x))
}
#include range to not get rid of any columns
obamahealth[,2:220]<-lapply(obamahealth[,2:220],normfunc)
numobama<-obamahealth[,2:220]
ObamaCorMat <- cor(numobama, method = "pearson")  ### only numeric vars
highlyCorrelated <- findCorrelation(ObamaCorMat, cutoff=0.75)
obamahealth <- obamahealth[,-highlyCorrelated]
rownames(obamahealth)<-NULL# to reset index

```



```{r backward selection}

#impute missing values to prevent errors in feature selection
#normalize the data
#do correlation analysis first
#make sure to only run models on numeric values
set.seed(1)
full <- lm(Class~.,data=obamahealth)
stepB <- stepAIC(full, direction= "backward", trace=FALSE)
summary(stepB)

```



```{r}
selectedobama<-obamahealth[ ,c(10,14,22,24,27,56,66,85,94,100)]
#rename class labels for proper glm
selectedobama$Class<-ifelse(selectedobama$Class==1,"inc","dec")
selectedobama$Class<-as.factor(selectedobama$Class)

```

```{r}
set.seed(1)
sampletrain <- sample(nrow(selectedobama), floor(nrow(selectedobama)*0.7))
selobtrain <- selectedobama[sampletrain,]#training set with labels
selobtest <- selectedobama[-sampletrain,]#test set with labels
selobtrainlabel<-selobtrain[,10]
selobtestlabel<-selobtest[,10]
selobtest<-selobtest[,-10]#test set no labels
```

```{r}
library(caretEnsemble)
set.seed(1)
control <- trainControl(method="repeatedcv", number = 10, repeats=5, savePredictions="final", classProbs = TRUE)
algorithmList <- c('rf', 'svmPoly', 'glm', 'naive_bayes')
set.seed(1)
models_selob <- caretList(Class~., data=selobtrain, trControl=control, methodList=algorithmList)
set.seed(1)
results_selob <- resamples(models_selob)
summary(results_selob)
dotplot(results_selob)  
modelCor(results_selob)

```

```{r}
#no model is very correlated but naive bayes has low accuracy
library(caretEnsemble)
set.seed(1)
control2 <- trainControl(method="repeatedcv", number = 10, repeats=5, savePredictions="final",classProbs = TRUE)
algorithmList2 <- c('rf', 'glm', 'svmPoly')
set.seed(1)
models_selob2 <- caretList(Class~., data=selobtrain, trControl=control2, methodList=algorithmList2)
set.seed(1)
results_selob2 <- resamples(models_selob2)
summary(results_selob2)
dotplot(results_selob2)  
modelCor(results_selob2)
```

```{r}
set.seed(1)
stackControl <- trainControl(method="repeatedcv", number=10, repeats=5, savePredictions=TRUE, classProbs=TRUE)
set.seed(1)
selobstack.rf <- caretStack(models_selob2, method="rf", metric="Accuracy", trControl=stackControl)
print(selobstack.rf)  ##ensemble model
```

```{r}
stack.pred <- predict(selobstack.rf , selobtest)
cf_ensemble_selob <- confusionMatrix(as.factor(stack.pred), as.factor(selobtestlabel) , positive="inc", mode = "everything")
print(cf_ensemble_selob)
```

# obama train, trump test 

```{r}
#first run feature selection
trumphealth<-trumphealthcare[,-c(93,112,114,223)]

```

```{r impute missing values}
for(i in 1:ncol(trumphealth)) {
  trumphealth[ , i][is.na(trumphealth[ , i])] <- mean(trumphealth[ , i], na.rm = TRUE)
}
#label the dataset in the data cleaning rmd
#now only pharmaceutical on the stock exchanges are included
notpharma<-c(which(trumphealth$naslabel==0 & trumphealth$nyselabel==0))
trumphealth<-trumphealth[-notpharma,]
sum(is.na(trumphealth))
```

```{r}
#normalize dataset
trumphealth[,2:220]<-lapply(trumphealth[,2:220],normfunc)

selectedtrumpfromob<-trumphealth[ ,c(41,51,72,76,84,155,168,205,215,221)]

selectedtrumpfromob$Class<-ifelse(selectedtrumpfromob$Class==1,"inc","dec")
selectedtrumpfromob$Class<-as.factor(selectedtrumpfromob$Class)
rownames(selectedtrumpfromob)<-NULL #reset index

```

```{r trump features from obama}
set.seed(1)
#sampletrain <- sample(nrow(selectedobama), floor(nrow(selectedobama)*0.7))
fullobtrain <- selectedobama[,]#training set with labels
fulltrumptest <- selectedtrumpfromob[,]#test set with labels
fullobamatrainlabel<-fullobtrain[,10]
fulltrumptestlabel<-fulltrumptest[,10]
fulltrumptest<-fulltrumptest[,-10]#test set no labels
```

```{r 70%obama model with fulltrump test}
OTstack.pred <- predict(selobstack.rf , fulltrumptest)
cf_ensemble_OT <- confusionMatrix(as.factor(OTstack.pred), as.factor(fulltrumptestlabel) , positive="inc", mode = "everything")
print(cf_ensemble_OT)
```

# Ensemble Model for training all of obama and only testing with trump, not used in report

```{r trump features from obama}
library(caretEnsemble)
set.seed(1)
control <- trainControl(method="repeatedcv", number = 5, repeats=3, savePredictions="final", classProbs = TRUE)
algorithmList <- c('rf', 'svmPoly', 'glm', 'naive_bayes')
set.seed(1)
models_fullob <- caretList(Class~., data=fullobtrain, trControl=control, methodList=algorithmList)
set.seed(1)
results_fullob <- resamples(models_fullob)
summary(results_fullob)
dotplot(results_fullob)  
modelCor(results_fullob)
#removing random forest because of NA and overfitting
```

```{r trump features from obama}
library(caretEnsemble)
set.seed(1)
control2 <- trainControl(method="repeatedcv", number = 5, repeats=3, savePredictions="final",classProbs = TRUE)
algorithmList2 <- c('svmPoly', 'glm', 'naive_bayes')
set.seed(1)
models_fullob2 <- caretList(Class~., data=fullobtrain, trControl=control, methodList=algorithmList2)
set.seed(1)
results_fullob2 <- resamples(models_fullob2)
summary(results_fullob2)
dotplot(results_fullob2)  
modelCor(results_fullob2)
```

```{r}
#second ensemble wasn't run because the models weren't corelated, changed models_fullob2 to models_fullob
set.seed(1)
stackControl <- trainControl(method="repeatedcv", number=10, repeats=3, savePredictions=TRUE, classProbs=TRUE)
set.seed(1)
fullobstack.rf <- caretStack(models_fullob, method="rf", metric="Accuracy", trControl=stackControl)
print(fullobstack.rf)  ##ensemble model
```

```{r}
fullobstack.pred <- predict(fullobstack.rf , fulltrumptest)
cf_ensemble_fullob <- confusionMatrix(as.factor(fullobstack.pred), as.factor(fulltrumptestlabel) , positive="inc", mode = "everything")
print(cf_ensemble_fullob)
```

# only trump ensemble model, 

```{r correlation and normalization}

trumphealth<-trumphealthcare[,-c(93,112,114,223)]

for(i in 1:ncol(trumphealth)) {
  trumphealth[ , i][is.na(trumphealth[ , i])] <- mean(trumphealth[ , i], na.rm = TRUE)
}
#log everything to normalize dataset, add high constant 
normfunc<-function(x){
  (x-min(x))/(max(x)-min(x))
}
#label the dataset in the data cleaning rmd
#now only pharmaceutical on the stock exchanges are included
notpharma<-c(which(trumphealth$naslabel==0 & trumphealth$nyselabel==0))
trumphealth<-trumphealth[-notpharma,]
#sum(is.na(obamahealth))
#include range to not get rid of any columns
trumphealth[,2:220]<-lapply(trumphealth[,2:220],normfunc)
numtrump<-trumphealth[,2:220]

TrumpCorMat <- cor(numtrump, method = "pearson")  ### only numeric vars
TrumpCorMat
highlyCorrelated <- findCorrelation(TrumpCorMat, cutoff=0.75)
trumphealth <- trumphealth[,-highlyCorrelated]
rownames(trumphealth)<-NULL# to reset index
# trump era has more correlated variables. non correlated left 100

```


```{r backward selection}

#impute missing values to prevent errors in feature selection
#normalize the data
#do correlation analysis first
#make sure to only run models on numeric values
fulltrump <- lm(Class~.,data=trumphealth)
trumpstepB <- stepAIC(fulltrump, direction= "backward", trace=FALSE)
summary(trumpstepB)

```

```{r}
selectedtrump<-trumphealth[ ,c(3,6,10,14,15,16,24,29,95,105,110,111,112,113)]
#rename class labels for proper glm
selectedtrump$Class<-ifelse(selectedtrump$Class==1,"inc","dec")
selectedobama$Class<-as.factor(selectedobama$Class)

```

```{r}
set.seed(1)
sampletrain <- sample(nrow(selectedobama), floor(nrow(selectedobama)*0.7))
selobtrain <- selectedobama[sampletrain,]#training set with labels
selobtest <- selectedobama[-sampletrain,]#test set with labels
selobtrainlabel<-selobtrain[,12]
selobtestlabel<-selobtest[,12]
selobtest<-selobtest[,-12]#test set no labels
```

```{r}
library(caretEnsemble)
set.seed(1)
control <- trainControl(method="repeatedcv", number = 5, repeats=3, savePredictions="final", classProbs = TRUE)
algorithmList <- c('rf', 'svmPoly', 'glm', 'naive_bayes')
set.seed(1)
models_selob <- caretList(Class~., data=selobtrain, trControl=control, methodList=algorithmList)
set.seed(1)
results_selob <- resamples(models_selob)
summary(results_selob)
dotplot(results_selob)  
modelCor(results_selob)

```

```{r}
library(caretEnsemble)
set.seed(1)
control2 <- trainControl(method="repeatedcv", number = 5, repeats=3, savePredictions="final",classProbs = TRUE)
algorithmList2 <- c('rf', 'glm', 'svmPoly')
set.seed(1)
models_selob2 <- caretList(Class~., data=selobtrain, trControl=control, methodList=algorithmList2)
set.seed(1)
results_selob2 <- resamples(models_selob2)
summary(results_selob2)
dotplot(results_selob2)  
modelCor(results_selob2)
```

```{r}
set.seed(1)
stackControl <- trainControl(method="repeatedcv", number=10, repeats=3, savePredictions=TRUE, classProbs=TRUE)
set.seed(1)
selobstack.rf <- caretStack(models_selob2, method="rf", metric="Accuracy", trControl=stackControl)
print(selobstack.rf)  ##ensemble model
```

```{r}
stack.pred <- predict(selobstack.rf , selobtest)
cf_ensemble_selob <- confusionMatrix(as.factor(stack.pred), as.factor(selobtestlabel) , positive="inc", mode = "everything")
print(cf_ensemble_selob)
```

