---
title: "Group 1 code"
author: "Group 1"
date: "12/9/2020"
output: pdf_document
---

#Question 1: Given today’s stock and pharmaceutical sales data, what is the best stock value to buy tomorrow?

## data cleaning

### Prepare Data for Merging
### 1. Load Data

```{r}
library(janitor)
library(tidyverse)
library(dplyr)
```


```{r}
# Load Stock Market Data of Pharmaceutical Companies
stock_market <- read_excel("Stock_Market_Compiled.xlsx")
# Load Daily Pharmaceutical Sales
daily_sales <- read_csv("salesdaily.csv")
```

### 2. Prepare Data and Merge
```{r}
# Average all pharmaceutical companies' stock values by date
stock_market_by_date <-
  stock_market %>%
  clean_names() %>% 
  group_by(date) %>% 
  summarise_if(.predicate = function(x) is.numeric(x),
               .funs = funs(mean="mean")) %>% 
  mutate(date = as.Date(date))
```

```{r}
# Reformat date to match stock market data for merging
daily_sales_date <-
  daily_sales  %>%
  clean_names() %>% 
  separate(datum, c("month", "date", "year"), "/") %>% 
  mutate(datum = paste(year, month, date, sep = "-"), 
         datum = as.Date(datum)) %>% 
  filter(!(year=="2019")) %>% # Include only 2014-2019 data
  select(c(datum, m01ab:weekday_name)) %>% 
  rename("date" = datum)
```

```{r}
# Merge all columns of stock market data and daily sales
daily_sales_stock <- full_join(daily_sales_date, stock_market_by_date, by = c("date"))
```

```{r}
# Write data to share with teammates
write.csv(daily_sales_stock, "daily_sales_stock.csv", row.names = F)
```

### 3. Prepare Merged Data
```{r}
dss_renamed <-
  daily_sales_stock %>% 
  rename_at(.vars = 2:9, ~c("Med4RheumArth","Med4OstArth", "Aspirin",
                            "Ibuprofen", "Med4Tension", "Med4Sleep", 
                            "Meds4Asthma", "Meds4Allergy"))
```

```{r}
library(tidyverse)
library(RCurl) # getURL 
library(MASS) # stepwise regression
library(leaps) # all subsets regression
library(corrplot)
library(caret)
library(FNN)
library(mlbench)
library(rcompanion)
library(e1071) 
```


#Import Merged Dataset and change column names of drugs (pharma data)
```{r}
data<-read.csv("daily_sales_stock.csv")
data<-drop_na(data)
colnames(data)[3:10]<-c("Med4RheumArth","Med4OstArth", "Aspirin","Ibuprofen", "Med4Tension", "Med4Sleep", "Meds4Asthma", "Meds4Allergy")
str(data)
```

#creating the column/variable low_price_next_day (Low Price of stock of next day) as our dependent variable
```{r}
shift <- function(x, n){
  c(x[-(seq(n))], rep(NA, n))}

data$low_price_next_day<-shift(data$low_mean,1)
data<-drop_na(data)
```

#Explore Dependent (Target) Varibale: low_price_next_day (Low Price of stock of next day)
```{r}
ggplot(data=data)+
  geom_boxplot(mapping = aes(x = low_price_next_day), color="#00AFBB") +
  coord_flip()+
  xlab("Low Stock Price of Next Day") +
  ggtitle("Distribution of Low Stock Price (2014-2018)") +
  theme(plot.title = element_text(size=12, face = "bold"))

plotNormalHistogram(data$low_price_next_day)

paste("Skewness of DV (next day low price):" ,skewness(data$low_price_next_day))


ggplot(data = data, aes(x = as.Date(date), y = low_price_next_day))+
  geom_line(color = "#00AFBB")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (2014-2018)") +
  theme(plot.title = element_text(size=12, face = "bold"))
```
The target variable is fairly symmetrical/normally distributed.


###Data Preprocessing (of IVs)
###Structure of Dataset
```{r}
data<-data[,-c(1,11)] #drop index(X)  and hour columns
str(data)
```

### Correlation Analysis
###Identify highly correlated "numeric" variables 
```{r}
###draw correlation matrix of the numeric independent variables only
num_data <- data[,-c(1,10,17)] ### remove date and weekday_name as well as low_price_next_day because it is our dependent variable
correlationMatrix <- cor(num_data, method = "pearson") 
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(correlationMatrix, method="color", col=col(200),   
         type="upper", order="hclust", 
         tl.col="black", tl.srt=45, tl.cex= 0.7, #Text label color and rotation
         # Combine with significance
         sig.level = 0.01, 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
)           

```

###Remove highly correlated numeric IVs (0.6)
```{r}
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.6)
noncor <- num_data[,-highlyCorrelated]  #keep only those not highly 
correlationMatrix2 <- cor(noncor, method = "pearson")  ### only numeric vars
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(correlationMatrix2, method="ellipse", col=col(200),   
         type="upper", order="hclust", 
         tl.col="black", tl.srt=45, tl.cex= 0.7, #Text label color and rotation
         # Combine with significance
         sig.level = 0.01, 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
)
```
###Normalizing numeric IVs
```{r}
normalize <- function(x) {
               return ((x - min(x)) / (max(x) - min(x))) }

noncor_n <- as.data.frame(lapply(noncor, normalize))

data_n<-cbind(noncor_n,data[,c(1,10,17)])
```


###Features Selection by Backward Elimination
```{r}
full <- lm(low_price_next_day~.,data=data_n[,-11])
stepB <- stepAIC(full, direction= "backward", trace=TRUE)
summary(stepB)
```
Significant IVs are: Med4RheumArth, Aspirin, Ibuprofen, Med4Tension, Meds4Asthma, Meds4Allergy, adj_close_mean, and volume_mean. Hence will drop Med4OstArth, Med4Sleep, Meds4Allergy, and weekday_name

###Dropping non-significant IVs (but still keeping date)
```{r}
data_n<-data_n[,-c(2,6,8,12)]
colnames(data_n)
```



##Code for 8 Machine Learning Models nd Auto Arima



###Linear Regression Model
### Model Assumptions
```{r}
LRmodel1 <- lm(low_price_next_day ~., data = train)
summary(LRmodel1)
plot(LRmodel1)
```
All four sssumptions of parametric linear regression were violated here. However, we decided to go against a transformation for interpretability and comparison (with other models) reasons.


###cross validation parameters
```{r}
ctrl <- trainControl(method="repeatedcv", number =5, repeats=3)  
```

###Model and distribution of errors
```{r}
set.seed(123)
LRmodel <- train(low_price_next_day ~ ., data= train, method="lm", trControl = ctrl)


test_predLR <- predict(LRmodel, test)
errors <- test_predLR - test_labels
summary(LRmodel)
hist(errors)
```


###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_labels - test_predLR)^2))

rel_change <- abs(errors) / test_labels
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```

###Creating the time series dataset for low stock price (of next day)
```{r}
data_ts<-data_n[,c(8,9)]
data_ts$date <- as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts[1:(cut-1),]
test_ts<-data_ts[cut:nrow(data_n),]
```

###Plot
```{r}
test_ts$preds<- test_predLR
ggplot(data = train_ts, aes(x = as.Date(date), y = low_price_next_day,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = low_price_next_day ,color="dodgerblue3"))+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "dodgerblue3", "orangered3"),
                       labels = c("Training", "Test", "Prediction"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): Linear Regression") +
  theme(plot.title = element_text(size=12, face = "bold"))
```

###KNN Model
```{r}
train_labels <- train[,8]      ##DV in the training set
test_labels <- test_labels     ##DV in the test set
train_set <- train[,-8]        ##Only keep IV in the training set
test_set <- test               ##Only keep IV in the test set
```

###Best value of K
```{r} 
x <- 0
for (i in 1:10)
{
KNNmodel <- knn.reg(train =train_set , test = test_set, y = train_labels , k = i)  
test_predKNN<- KNNmodel$pred
rmse <- sqrt(mean((test_labels-test_predKNN)^2))
x[i] <- rmse
}

plot(x, type="l", col="red")
bk<-which.min(x)
paste('Best K(by RMSE):', bk)
```

###Model and distribution of errors
```{r}
set.seed(123)
KNNmodel <- knn.reg(train =train_set , test = test_set, y = train_labels , k =bk)
test_predKNN <- KNNmodel$pred
errors <- test_predKNN - test_labels
hist(errors)
```

###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_labels-test_predKNN)^2))

rel_change <- abs(errors) / test_labels
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test_set)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test_test)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test_set)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test_set)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test_set)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```


###Re-creating the time series dataset for low stock price (of next day)
```{r}
data_ts<-data_n[,c(8,9)]
data_ts$date <- as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts[1:(cut-1),]
test_ts<-data_ts[cut:nrow(data_n),]
```

###Plot
```{r}
test_ts$preds<- test_predKNN
ggplot(data = train_ts, aes(x = as.Date(date), y = low_price_next_day,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = low_price_next_day ,color="dodgerblue3"))+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "dodgerblue3", "orangered3"),
                       labels = c("Training", "Test", "Prediction"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): KNN Regression") +
  theme(plot.title = element_text(size=12, face = "bold"))
```

###Random Forest Model
###Best number of trees
```{r}
set.seed(123)
library(randomForest)
m1 <- randomForest(formula = low_price_next_day ~ ., data = train)

plot(m1)
bntrees<-which.min(m1$mse)
paste("Best number of tree:", bntrees)
```

###Model and distribution of errors



```{r}
set.seed(123)
RFmodel <- train(low_price_next_day ~ ., data= train, method="rf", ntree=bntrees, trControl = ctrl)

test_predRF <- predict(RFmodel, test)
errors <- test_predRF - test_labels
hist(errors)
```

###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_labels-test_predRF)^2))

rel_change <- abs(errors) / test_labels
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```

###Re-creating the time series dataset for low stock price (of next day)
```{r}
data_ts<-data_n[,c(8,9)]
data_ts$date <- as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts[1:(cut-1),]
test_ts<-data_ts[cut:nrow(data_ts),]
```

###Plot
```{r}
test_ts$preds<- test_predRF
ggplot(data = train_ts, aes(x = as.Date(date), y = low_price_next_day,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = low_price_next_day ,color="dodgerblue3"))+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "dodgerblue3", "orangered3"),
                       labels = c("Training", "Test", "Prediction"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): Random Forest Regression") +
  theme(plot.title = element_text(size=12, face = "bold"))
```

###SVM Regression
###Model and distribution of errors
```{r}
set.seed(123)
SVMmodel <- train(low_price_next_day ~ ., data= train, method="svmPoly", trControl = ctrl)

test_predSVM <- predict(SVMmodel, test)
errors <- test_predSVM - test_labels
hist(errors)
```

###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_labels-test_predSVM)^2))

rel_change <- abs(errors) / test_labels
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```

###Re-creating the time series dataset for low stock price (of next day)
```{r}
data_ts<-data_n[,c(8,9)]
data_ts$date <- as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts[1:(cut-1),]
test_ts<-data_ts[cut:nrow(data_ts),]
```

###Plot
```{r}
test_ts$preds<- test_predSVM
ggplot(data = train_ts, aes(x = as.Date(date), y = low_price_next_day,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = low_price_next_day ,color="dodgerblue3"))+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "dodgerblue3", "orangered3"),
                       labels = c("Training", "Test", "Prediction"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): SVM Regression") +
  theme(plot.title = element_text(size=12, face = "bold"))
```

###Ensemble Model
```{r}
library(caretEnsemble)
control <- trainControl(method="repeatedcv", number = 5, repeats=3, savePredictions="final")
algorithmList <- c('rf', 'svmPoly', 'lm')
set.seed(123)
models_ens <- caretList(low_price_next_day~., data= train, trControl=control, methodList=algorithmList)
set.seed(123)
results <- resamples(models_ens)
summary(results)
dotplot(results)  
modelCor(results)
```
SVM is highly correlated with random forest(RF) and linear regression(LR/LM). Hence, we will keep only random forest and linear regression.

###only keep RF and LR/LM
```{r}
library(caretEnsemble)
control <- trainControl(method="repeatedcv", number = 5, repeats=3, savePredictions="final")
algorithmList <- c('rf', 'lm')
set.seed(123)
models_ens2 <- caretList(low_price_next_day~., data= train, trControl=control, methodList=algorithmList)
set.seed(123)
results2 <- resamples(models_ens2)
summary(results2)
dotplot(results2)  
modelCor(results2)
```

###Combine predictions of the final models (RF and LR/LM) used using stack with RF
```{r}
stackControl <- trainControl(method="repeatedcv", number=5, repeats=3, savePredictions=TRUE)
set.seed(123)
stack.rf <- caretStack(models_ens2, method="rf", metric="RMSE", trControl=stackControl)
print(stack.rf)  ##ensemble model
```

###predict stack.rf on the test dataset and plot its errors
```{r}
test_predEM <- predict(stack.rf , test)
errors <- test_predEM - test_labels
hist(errors)
```

###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_labels-test_predEM)^2))

rel_change <- abs(errors) / test_labels
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```

###Re-creating the time series dataset for low stock price (of next day)
```{r}
data_ts<-data_n[,c(8,9)]
data_ts$date <- as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts[1:(cut-1),]
test_ts<-data_ts[cut:nrow(data_ts),]
```

###Plot
```{r}
test_ts$preds<- test_predEM
ggplot(data = train_ts, aes(x = as.Date(date), y = low_price_next_day,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = low_price_next_day ,color="dodgerblue3"))+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "dodgerblue3", "orangered3"),
                       labels = c("Training", "Test", "Prediction"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): Ensemble Model (RF&LR)") +
  theme(plot.title = element_text(size=12, face = "bold"))
```

###Simple Moving Average (of last 250 low stock price)
###Re-creating the time series dataset for low stock price
```{r}
data_ts<-data[,c(1,13)]
data_ts$date = as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts[1:(cut-1),]
test_ts<-data_ts[cut:nrow(data_ts),]
```

###Model and distribution of errors
```{r}
index<-nrow(test_ts)
preds = 0
for (i in 1:index)
  {
  a = sum(train_ts$low_mean[(nrow(train_ts)-index+i):(nrow(train_ts))]) + sum(preds)
  b = a/index
  preds[i] <- b
}
errors <- preds - test_ts$low_mean
hist(errors)
```
###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_ts$low_mean - preds)^2))

rel_change <- abs(errors) / test_ts$low_mean
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test_ts)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test_ts)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```

###Plot
```{r}
test_ts$preds <- preds
ggplot(data = train_ts, aes(x = as.Date(date), y = low_mean,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = low_mean ,color="dodgerblue3"))+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds ,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "dodgerblue3", "orangered3"),
                       labels = c("Training", "Test", "Predictions"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): Moving Average Method") +
  theme(plot.title = element_text(size=12, face = "bold"))
```

###Auto-ARIMA
###Re-creating the time series dataset for low stock price
```{r}
data_ts<-data[,c(1,13)]
data_ts$date = as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts[1:(cut-1),]
test_ts<-data_ts[cut:nrow(data_ts),]
```

###Model and distribution of errors
```{r}
#install.packages("forecast")
library(forecast)
ARIMAmodel <- auto.arima(train_ts$low_mean,
                  max.p = 5,
                  max.q = 5,
                  max.P = 2,
                  max.Q = 2,
                  max.order = 5,
                  max.d = 2,
                  max.D = 1,
                  start.p = 2,
                  start.q = 2,
                  start.P = 1,
                  start.Q = 1,
                  stationary = FALSE,
                  seasonal = TRUE,
                  ic = c("aicc", "aic", "bic"),
                  stepwise = TRUE,
                  nmodels = 100,
                  trace = FALSE,
                  method = NULL,
                  truncate = NULL,
                  test = c("kpss", "adf", "pp"),
                  seasonal.test = c("seas", "ocsb", "hegy", "ch"),
                  allowdrift = TRUE,
                  lambda = NULL,
                  biasadj = FALSE,
                  parallel = FALSE)

test_predARIMA<-predict(ARIMAmodel,n.ahead = nrow(test_ts))
```


```{r}
test_predARIMA<- test_predARIMA$pred
errors <- test_predARIMA - test_ts$low_mean
hist(errors)
```

###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_ts$low_mean - test_predARIMA)^2))

rel_change <- abs(errors) / test_ts$low_mean
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test_ts)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test_ts)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```


###Plot
```{r}
test_ts$preds <- test_predARIMA
ggplot(data = train_ts, aes(x = as.Date(date), y = low_mean,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = low_mean ,color="dodgerblue3"))+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "dodgerblue3", "orangered3"),
                       labels = c("Training", "Test", "Prediction"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): ARIMA Method") +
  theme(plot.title = element_text(size=12, face = "bold"))
```

One might ask why we cannot just average out the low stock price of the previous 2, 3, 5 days and go with it rather than engage in a "complex ML process". Averaging out the low stock values of the past 5 previous days seems reasonable and fair since the stock market operates only on weekdays (except holidays). Let's try and see. 

###Average of Previous '5 Days' (of low stock price to predict next low stock price) Method
###Re-creating the time series dataset for low stock price
```{r}
data_ts<-data[,c(1,13)]
data_ts$date = as.Date(data_ts$date)
```

```{r}
train_ts<-data_ts
test_ts<-data_ts[6:nrow(data_ts),]
```

###Model and distribution of errors
```{r}
index<-nrow(test_ts)
preds = 0
for (i in 1:index)
  {
  a = sum(train_ts$low_mean[i:(i+4)])
  b = a/5
  preds[i] <- b
}
errors <- preds - test_ts$low_mean
hist(errors)
```

###RMSE and cases with less than 25%, 10%, 5%, and 1% error
```{r}
rmse <- sqrt(mean((test_ts$low_mean - preds)^2))

rel_change <- abs(errors) / test_ts$low_mean
pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test_ts)   ## gives the count of those who are true on the condition of rel_change<25%
##OR pred25 <- table(rel_change<0.25)["TRUE"] / nrow(test_ts)

pred10 <- sum((rel_change<0.10)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<10%

pred5 <- sum((rel_change<0.05)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<5%

pred1 <- sum((rel_change<0.01)=="TRUE")/nrow(test_ts)  ## gives the count of those who are true on the condition of rel_change<1%

paste("RMSE:", rmse)
paste("Pred(25%):", pred25)
paste("Pred(10%):", pred10)
paste("Pred(5%):", pred5)
paste("Pred(1%):", pred1)
```

###Plot
```{r}
test_ts$preds <- preds
ggplot(data = test_ts, aes(x = as.Date(date), y = low_mean,color = "aquamarine3"))+
  geom_line()+
  geom_line(data = test_ts, aes(x = as.Date(date), y = preds ,color="orangered3"))+
  scale_color_identity(name = " ",
                       breaks = c("aquamarine3", "orangered3"),
                       labels = c("Training/Test", "Predictions"),
                       guide = "legend")+
  xlab("Time")+
  ylab("Low Stock Price of Next Day")+
  ggtitle("Low Stock Price (Next Day): Average of Previous '5 Days'  Method") +
  theme(plot.title = element_text(size=12, face = "bold"))
```
Averaging out the low stock price registered on the five previous days seems to perform poorly when there is a sharp change.


# question 2
# question 2 part 1:  Are there changes in the types of drugs sold in the years after Trump was elected? 
## data cleaning  

```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(rcompanion)
library(MASS)# stepwise regression
library(randomForest) 
library(leaps) # all subsets regression
library(corrplot)
library(caret)
library(FNN)
library(mlbench)
library(lattice)
library(class)
library(gmodels)
library(mice)
```

```{r}
#run this first
# Pharma sales dataset
#do linear regression to check of correlation for between and within years
weeklypharma<-read.csv("salesweekly.csv")
colnames(weeklypharma)
colnames(weeklypharma)<-c("Date","Med4RheumArth","Med4OstArth","Aspirin","Ibuprofen",
                          "Med4Tension","Med4Sleep","Med4Asthma","Med4Allergy")
str(weeklypharma)
#convert to date time object
weeklypharma$Date<-mdy(weeklypharma$Date)

#Obama year sales
pharma201415<-weeklypharma
pharma201415 <- 
  pharma201415 %>% 
  filter(between(Date,as.Date("2014-01-01"), as.Date("2015-12-31"))) 

#Trump year sales
pharma201718<-weeklypharma
pharma201718 <- 
  pharma201718 %>% 
  filter(between(Date,as.Date("2017-01-01"), as.Date("2018-12-31")))

#election year sales
pharma2016<-weeklypharma
pharma2016 <- 
  pharma2016 %>% 
  filter(between(Date,as.Date("2016-01-01"), as.Date("2016-12-31")))
```

```{r obama data exploration and transformation for linear reg}
# first, load cleaned datasets from FinalProjectDataCleaning
obamapharma<-pharma201415[,-1]
correlationMatrix <- cor(obamapharma, method = "pearson")
correlationMatrix
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(correlationMatrix, method="color", col=col(200),type="upper", order="hclust")
# MEdicine are not highly correlated (all meds were 0.4 or under) with each other therefore each no med was dropped. 
colnames(pharma201415)
str(pharma201415)
library(e1071)                    # load e1071 
skewness(pharma201415$Med4Sleep)
#plotNormalHistogram(pharma201415$Meds4Allergy)
#no missing values in the phamadata
#med4rheumarth skewness is 0.1803
#med4ostarth skew is -0.025
#med4sleep skew is 0.897, rskew
#med4asthma skew is 0.97, r skew
#med4allergy skew is 0.9579, rskew
#med4tension is 0.793, rskew
#ibuprofen is 0.79731, rskew
#aspirin is 0.5888, right skewed
#6/8 variables are right skewed, log transformation is done
#("Date","Med4RheumArth","Med4OstArth"," Aspirin","Ibuprofen"," Med4Tension","Med4Sleep","Meds4Asthma","Meds4Allergy")

#log transformations required, add small coefficient (0.5) to prevent log of 0
#do them individually so that the dataframe isnt spliced
dolog<-function(x){
  log(x+1)
}
t201415<-pharma201415
t201415[2:9]<-lapply(t201415[2:9],dolog)#logtransformed

```

###Stepwise Feature Selection for obama era
Since none of them were correlated, hard to know which features are best predictors of independent variables
```{r feature selection for rheumatoid arthritis meds}
fullobamarh <- lm(Med4RheumArth~., data= t201415)  ## . means all the IVs
nullobamarh <- lm(Med4RheumArth~1,data=t201415)
stepobamarh <- stepAIC(nullobamarh, scope=list(lower=nullobamarh, upper=fullobamarh), direction= "both", trace=FALSE)
summary(stepobamarh) 
```

```{r feature selection for osteo arthritis meds}
fullobamaost <- lm(Med4OstArth~., data= t201415)  ## . means all the IVs
nullobamaost <- lm(Med4OstArth~1,data=t201415)
stepobamaost <- stepAIC(nullobamaost, scope=list(lower=nullobamaost, upper=fullobamaost), direction= "both", trace=FALSE)
summary(stepobamaost) 

```

```{r feature selection for aspirin}
fullobamaasp <- lm(Aspirin~., data= t201415)  ## . means all the IVs
nullobamaasp <- lm(Aspirin~1,data=t201415)
stepobamaasp <- stepAIC(nullobamaasp, scope=list(lower=nullobamaasp, upper=fullobamaasp), direction= "both", trace=FALSE)
summary(stepobamaasp) 
```

```{r feature selection for ibuprofen}
fullobamaib <- lm(Ibuprofen~., data= t201415)  ## . means all the IVs
nullobamaib <- lm(Ibuprofen~1,data=t201415)
stepobamaib <- stepAIC(nullobamaib, scope=list(lower=nullobamaib, upper=fullobamaib), direction= "both", trace=FALSE)
summary(stepobamaib) 

```

```{r feature selection for tension medicine}
fullobamaten <- lm(Med4Tension~., data= t201415)  ## . means all the IVs
nullobamaten <- lm(Med4Tension~1,data=t201415)
stepobamaten <- stepAIC(nullobamaten, scope=list(lower=nullobamaten, upper=fullobamaten), direction= "both", trace=FALSE)
summary(stepobamaten) 
```

```{r feature selection for sleep medicine}
fullobamaslp <- lm(Med4Sleep~., data= t201415)  ## . means all the IVs
nullobamaslp <- lm(Med4Sleep~1,data=t201415)
stepobamaslp <- stepAIC(nullobamaslp, scope=list(lower=nullobamaslp, upper=fullobamaslp), direction= "both", trace=FALSE)
summary(stepobamaslp) 
```

```{r feature selection for asthma medicine}
fullobamaast <- lm(Med4Asthma~., data= t201415)  ## . means all the IVs
nullobamaast <- lm(Med4Asthma~1,data=t201415)
stepobamaast <- stepAIC(nullobamaast, scope=list(lower=nullobamaast, upper=fullobamaast), direction= "both", trace=FALSE)
summary(stepobamaast) 
```

```{r feature selection for allergy medicine}
fullobamaall <- lm(Med4Allergy~., data= t201415)  ## . means all the IVs
nullobamaall <- lm(Med4Allergy~1,data=t201415)
stepobamaall <- stepAIC(nullobamaall, scope=list(lower=nullobamaall, upper=fullobamaall), direction= "both", trace=FALSE)
summary(stepobamaall) 
```

### Trump era medicines Stepwise Feature Selection
Since none of them were correlated, hard to know which features are best predictors of independent variables
```{r feature selection for rheumatoid arthritis meds}
fulltrumprh <- lm(Med4RheumArth~., data= t201718)  ## . means all the IVs
nulltrumprh <- lm(Med4RheumArth~1,data=t201718)
steptrumprh <- stepAIC(nulltrumprh, scope=list(lower=nulltrumprh, upper=fulltrumprh), direction= "both", trace=FALSE)
summary(steptrumprh) 
```

```{r feature selection for osteo arthritis meds}
fulltrumpost <- lm(Med4OstArth~., data= t201718)  ## . means all the IVs
nulltrumpost <- lm(Med4OstArth~1,data=t201718)
steptrumpost <- stepAIC(nulltrumpost, scope=list(lower=nulltrumpost, upper=fulltrumpost), direction= "both", trace=FALSE)
summary(steptrumpost) 

```

```{r feature selection for aspirin}
fulltrumpasp <- lm(Aspirin~., data= t201718)  ## . means all the IVs
nulltrumpasp <- lm(Aspirin~1,data=t201718)
steptrumpasp <- stepAIC(nulltrumpasp, scope=list(lower=nulltrumpasp, upper=fulltrumpasp), direction= "both", trace=FALSE)
summary(steptrumpasp) 
```

```{r feature selection for ibuprofen}
fulltrumpib <- lm(Ibuprofen~., data= t201718)  ## . means all the IVs
nulltrumpib <- lm(Ibuprofen~1,data=t201718)
steptrumpib <- stepAIC(nulltrumpib, scope=list(lower=nulltrumpib, upper=fulltrumpib), direction= "both", trace=FALSE)
summary(steptrumpib) 

```

```{r feature selection for tension medicine}
fulltrumpten <- lm(Med4Tension~., data= t201718)  ## . means all the IVs
nulltrumpten <- lm(Med4Tension~1,data=t201718)
steptrumpten <- stepAIC(nullobamaten, scope=list(lower=nulltrumpten, upper=fulltrumpten), direction= "both", trace=FALSE)
summary(steptrumpten) 
```

```{r feature selection for sleep medicine}
fulltrumpslp <- lm(Med4Sleep~., data= t201718)  ## . means all the IVs
nulltrumpslp <- lm(Med4Sleep~1,data=t201718)
steptrumpslp <- stepAIC(nulltrumpslp, scope=list(lower=nulltrumpslp, upper=fulltrumpslp), direction= "both", trace=FALSE)
summary(steptrumpslp) 
```

```{r feature selection for asthma medicine}
fulltrumpast <- lm(Med4Asthma~., data= t201718)  ## . means all the IVs
nulltrumpast <- lm(Med4Asthma~1,data=t201718)
steptrumpast <- stepAIC(nulltrumpast, scope=list(lower=nulltrumpast, upper=fulltrumpast), direction= "both", trace=FALSE)
summary(steptrumpast) 
```

```{r feature selection for allergy medicine}
fulltrumpall <- lm(Med4Allergy~., data= t201718)  ## . means all the IVs
nulltrumpall <- lm(Med4Allergy~1,data=t201718)
steptrumpall <- stepAIC(nulltrumpall, scope=list(lower=nulltrumpall, upper=fulltrumpall), direction= "both", trace=FALSE)
summary(steptrumpall) 
```

NOTES: stepwise regression was done for weekly on the time periods 201415 and 201718. for the 201415 period, 5/8 meds had date as significant but only 1/8 meds had date as signifcant for trump era.

### Question 2 part 1 Training and testing 2014-2015 models to make sure model gives back good metrics before testing on other data

```{r traintest split }
#linear reg train test split
set.seed(111)
rn_train <- sample(nrow(t201415), floor(nrow(t201415)*0.7))
obamatrain70<-t201415[rn_train,]
obamatest<-t201415[-rn_train,]


```

Using Linear Regression to train 70% 201415 data and test 30% on 20142015
rheumArth, Aspirin,Ibuprofen,Sleep Med and Allergy Med had date be a significant feature so only those 5 meds were used for training and test.

```{r linear model for rheum arthritis}
ObamaRheumModel<-lm(Med4RheumArth~Date, data=obamatrain70) 
ObamaRheumPrediction <- predict(ObamaRheumModel, newdata =obamatest) 
ObamaRheumErrors <- ObamaRheumPrediction - obamatest$Med4RheumArth
hist(ObamaRheumErrors)
ObamaRheumRmse <- sqrt(mean((obamatest$Med4RheumArth - ObamaRheumPrediction)^2))
ObamaRheumRel_change <- abs(ObamaRheumErrors) / obamatest$Med4RheumArth
ObamaRheumPred10 <- table(ObamaRheumRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaRheumPred25 <- table(ObamaRheumRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Rheumatoid Arthritis Meds")
paste("RMSE:", ObamaRheumRmse)
paste("PRED(10):", round(ObamaRheumPred10,2))
paste("PRED(25):", round(ObamaRheumPred25,2))
paste("Summary of Prediction")
summary(ObamaRheumPrediction)


#88%of values are within 10% of the actual values, 100% of values are within 25% of actualvalues and there is a very low RMSE(0.178) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Aspirin}
ObamaAspModel<-lm(Aspirin~Date, data=obamatrain70) 
ObamaAspPrediction <- predict(ObamaAspModel, newdata =obamatest) 
ObamaAspErrors <- ObamaAspPrediction - obamatest$Aspirin
hist(ObamaAspErrors)
ObamaAspRmse <- sqrt(mean((obamatest$Aspirin - ObamaAspPrediction)^2))
ObamaAspRel_change <- abs(ObamaAspErrors) / obamatest$Aspirin

ObamaAspPred10 <- table(ObamaAspRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaAspPred25 <- table(ObamaAspRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Aspirin")
paste("RMSE:", ObamaAspRmse)
paste("PRED(10):", round(ObamaAspPred10,2))
paste("PRED(25):", round(ObamaAspPred25,2))
paste("Summary of Prediction")
summary(ObamaAspPrediction)

#within 201415
#88% of values are within 10% of the actual values and 100% of values are within 25% of actual values there is a very low RMSE(0.243) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Ibuprofen}

ObamaIbModel<-lm(Ibuprofen~Date, data=obamatrain70) 
ObamaIbPrediction <- predict(ObamaIbModel, newdata =obamatest) 
ObamaIbErrors <- ObamaIbPrediction - obamatest$Ibuprofen
hist(ObamaIbErrors)
ObamaIbRmse <- sqrt(mean((obamatest$Ibuprofen - ObamaIbPrediction)^2))
ObamaIbRel_change <- abs(ObamaIbErrors) / obamatest$Ibuprofen
ObamaIbPred10 <- table(ObamaIbRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaIbPred25 <- table(ObamaIbRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Ibuprofen")
paste("RMSE:", ObamaIbRmse)
paste("PRED(10):", round(ObamaIbPred10,2))
paste("PRED(25):", round(ObamaIbPred25,2))
paste("Summary of Prediction")
summary(ObamaIbPrediction)
#within201415

#88%of values are within 10% of the actual values and 100% of values fall within 25% of actual values (match 100%) and there is a very low RMSE(0.283) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for sleep med}
ObamaSlpModel<-lm(Med4Sleep~Date, data=obamatrain70) 
ObamaSlpPrediction <- predict(ObamaSlpModel, newdata =obamatest) 
ObamaSlpErrors <- ObamaSlpPrediction - obamatest$Med4Sleep
hist(ObamaSlpErrors)
ObamaSlpRmse <- sqrt(mean((obamatest$Med4Sleep - ObamaSlpPrediction)^2))
ObamaSlpRel_change <- abs(ObamaSlpErrors) / obamatest$Med4Sleep
ObamaSlpPred10 <- table(ObamaSlpRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaSlpPred25 <- table(ObamaSlpRel_change<0.25)["TRUE"] / nrow(obamatest)

paste("2014-2015 model for Sleep Meds")
paste("RMSE:", ObamaSlpRmse)
paste("PRED(10):", round(ObamaSlpPred10,2))
paste("PRED(25):", round(ObamaSlpPred25,2))
paste("Summary of Prediction")
summary(ObamaSlpPrediction)

#22%of values are within 10% of the actual values, 66% fall within 25% of the actual values and there is a very moderate RMSE(0.65) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for allergymed}

ObamaAllModel<-lm(Med4Allergy~Date, data=obamatrain70) 
ObamaAllPrediction <- predict(ObamaAllModel, newdata =obamatest) 
ObamaAllErrors <- ObamaAllPrediction - obamatest$Med4Allergy
hist(ObamaAllErrors)
ObamaAllRmse <- sqrt(mean((obamatest$Med4Allergy - ObamaAllPrediction)^2))
ObamaAllRel_change <- abs(ObamaAllErrors) / obamatest$Med4Allergy
ObamaAllPred10 <- table(ObamaAllRel_change<0.10)["TRUE"] / nrow(obamatest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
ObamaAllPred25 <- table(ObamaAllRel_change<0.25)["TRUE"] / nrow(obamatest)
paste("2014-2015 model for Allergy Meds")
paste("RMSE:", ObamaAllRmse)
paste("PRED(10):", round(ObamaAllPred10,2))
paste("PRED(25):", round(ObamaAllPred25,2))
paste("Summary of Prediction")
summary(ObamaAllPrediction)
#41%of values are within 10% of the actual values, 81.25% are within 25% of actual values and there is a very low RMSE(0.54) which may indicate that the sales of medicines follows a certain pattern.
```


###Obama train, trump test model
###Obama train, trump test

Using Linear Regression to train 201415 data and test on 201718.
rheumArth, Aspirin,Ibuprofen,Sleep Med and Allergy Med had date be a significant feature so only those 5 meds were used for training and test.


```{r traintest split }
#linear reg train test split
#rn_train <- sample(nrow(t201415), floor(nrow(t201415)*0.7))
#obamatrain<-t20142015[rn_train,]
#obamatest<-t20142015[-rn_train,]

obamatrain<-t201415
trumptest<-t201718[-1,]#to keep the number of observations the same
# to reset index rownames(trumptest)<-NULL

```

```{r linear model for rheum arthritis}
OTRheumModel<-lm(Med4RheumArth~Date, data=obamatrain) 
OTRheumPrediction <- predict(OTRheumModel, newdata =trumptest) 
OTRheumErrors <- OTRheumPrediction - trumptest$Med4RheumArth
hist(OTRheumErrors)
OTRheumRmse <- sqrt(mean((trumptest$Med4RheumArth - OTRheumPrediction)^2))
OTRheumRel_change <- abs(OTRheumErrors) / trumptest$Med4RheumArth
OTRheumPred10 <- table(OTRheumRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTRheumPred25 <- table(OTRheumRel_change<0.25)["TRUE"] / nrow(trumptest)
paste("TrumpTest: Rheumatoid Athritis Med")
paste("RMSE:", OTRheumRmse)
paste("PRED(10):", round(OTRheumPred10,2))
paste("PRED(25):", round(OTRheumPred25,2))
paste("Summary of Prediction")
summary(OTRheumPrediction)
plot(OTRheumErrors)
plot(fitted(OTRheumModel), resid(OTRheumModel))
#91%of values are within 10% of the actual values, 100% of values are within 25% of actualvalues and there is a very low RMSE(0.178) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Aspirin}
OTAspModel<-lm(Aspirin~Date, data=obamatrain) 
OTAspPrediction <- predict(OTAspModel, newdata=trumptest) 
OTAspErrors <- OTAspPrediction - trumptest$Aspirin
hist(OTAspErrors)
OTAspRmse <- sqrt(mean((trumptest$Aspirin - OTAspPrediction)^2))
OTAspRel_change <- abs(OTAspErrors) / trumptest$Aspirin

OTAspPred10 <- table(OTAspRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTAspPred25 <- table(OTAspRel_change<0.25)["TRUE"] / nrow(trumptest)
paste("TrumpTest: Aspirin")
paste("RMSE:", OTAspRmse)
paste("PRED(10):", round(OTAspPred10,2))
paste("PRED(25):", round(OTAspPred25,2))
paste("Summary of Prediction")
summary(OTAspPrediction)

#60% of values are within 10% of the actual values and 91% of values are within 25% of actual values there is a very low RMSE(0.243) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for Ibuprofen}

OTIbModel<-lm(Ibuprofen~Date, data=obamatrain) 
OTIbPrediction <- predict(OTIbModel, newdata =trumptest) 
OTIbErrors <- OTIbPrediction - trumptest$Ibuprofen
hist(OTIbErrors)
OTIbRmse <- sqrt(mean((trumptest$Ibuprofen - OTIbPrediction)^2))
OTIbRel_change <- abs(OTIbErrors) / trumptest$Ibuprofen
OTIbPred10 <- table(OTIbRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTIbPred25 <- table(OTIbRel_change<0.25)["TRUE"] / nrow(trumptest)

paste("TrumpTest: Ibuprofen")
paste("RMSE:", OTIbRmse)
paste("PRED(10):", round(OTIbPred10,2))
paste("PRED(25):", round(OTIbPred25,2))
paste("Summary of Prediction")
summary(OTIbPrediction)


# 26% of predicted values fall within 25% of actual values . 91%of values are within 25% of the actual values and there is a very low RMSE(0.283) which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for sleep med}

OTSlpModel<-lm(Med4Sleep~Date, data=obamatrain) 
OTSlpPrediction <- predict(OTSlpModel, newdata=trumptest) 
OTSlpErrors <- OTSlpPrediction - trumptest$Med4Sleep
hist(OTSlpErrors)
OTSlpRmse <- sqrt(mean((trumptest$Med4Sleep - OTSlpPrediction)^2))
OTSlpRel_change <- abs(OTSlpErrors) / trumptest$Med4Sleep
OTSlpPred10 <- table(OTSlpRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTSlpPred25 <- table(OTSlpRel_change<0.25)["TRUE"] / nrow(trumptest)

paste("TrumpTest: Sleep Medicine")
paste("RMSE:", OTSlpRmse)
paste("PRED(10):", round(OTSlpPred10,2))
paste("PRED(25):", round(OTSlpPred25,2))
paste("Summary of Prediction")
summary(OTSlpPrediction)

#None of the values are within 10% of the actual values, 2% fall within 25% of the actual values and there is an RMSE 1.25 which is very high which may indicate that the sales of medicines follows a certain pattern.
```

```{r linear model for allergymed}

OTAllModel<-lm(Med4Allergy~Date, data=obamatrain) 
OTAllPrediction <- predict(OTAllModel, newdata =trumptest) 
OTAllErrors <- OTAllPrediction - trumptest$Med4Allergy
hist(OTAllErrors)
OTAllRmse <- sqrt(mean((trumptest$Med4Allergy - OTAllPrediction)^2))
OTAllRel_change <- abs(OTAllErrors) / trumptest$Med4Allergy
OTAllPred10 <- table(OTAllRel_change<0.10)["TRUE"] / nrow(trumptest)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
OTAllPred25 <- table(OTAllRel_change<0.25)["TRUE"] / nrow(trumptest)

paste("TrumpTest: Allergy Medicine")
paste("RMSE:", OTAllRmse)
paste("PRED(10):", round(OTAllPred10,2))
paste("PRED(25):", round(OTAllPred25,2))
paste("Summary of Predictions")
summary(OTAllPrediction)

#41%of values are within 10% of the actual values, 81.25% are within 25% of actual values and there is a very low RMSE(0.54) which may indicate that the sales of medicines follows a certain pattern.
```



# question 2 part 2: Additionally, are there general changes in the overall financial standing of the pharmaceutical industry between 2014 and 2018?  
## Regression and Classification methods were used for this question


### data cleaning for Muliple linear regression and KNN


#### Load Financial Indicators dataset
```{r}
fin2014 <- read.csv("/Users/aselkushkeyeva/Desktop/Year2/Fall2020/INF2167AppliedDataScienceUsingR/FinalProject/200FinIndicatorsUSstocks/2014_Financial_Data.csv")
fin2015 <- read.csv("/Users/aselkushkeyeva/Desktop/Year2/Fall2020/INF2167AppliedDataScienceUsingR/FinalProject/200FinIndicatorsUSstocks/2015_Financial_Data.csv")
fin2016 <- read.csv("/Users/aselkushkeyeva/Desktop/Year2/Fall2020/INF2167AppliedDataScienceUsingR/FinalProject/200FinIndicatorsUSstocks/2016_Financial_Data.csv")
fin2017 <- read.csv("/Users/aselkushkeyeva/Desktop/Year2/Fall2020/INF2167AppliedDataScienceUsingR/FinalProject/200FinIndicatorsUSstocks/2017_Financial_Data.csv")
fin2018 <- read.csv("/Users/aselkushkeyeva/Desktop/Year2/Fall2020/INF2167AppliedDataScienceUsingR/FinalProject/200FinIndicatorsUSstocks/2018_Financial_Data.csv")
```

#### Merge the datasets
```{r}
#rename the non-matching column to merge the datasets
fin2014$X201X.PRICE.VAR.... <- fin2014$X2015.PRICE.VAR.... 
fin2015$X201X.PRICE.VAR.... <- fin2015$X2016.PRICE.VAR....
fin2016$X201X.PRICE.VAR.... <- fin2016$X2017.PRICE.VAR....
fin2017$X201X.PRICE.VAR.... <- fin2017$X2018.PRICE.VAR....
fin2018$X201X.PRICE.VAR.... <- fin2018$X2019.PRICE.VAR....

#Remove the original non-matching columns
fin2014 <- fin2014[, -224]
fin2015 <- fin2015[, -224]
fin2016 <- fin2016[, -224]
fin2017 <- fin2017[, -224]
fin2018 <- fin2018[, -224]

# Add Year column in each dataset
fin2014$year <- c(2014)
fin2015$year <- c(2015)
fin2016$year <- c(2016)
fin2017$year <- c(2017)
fin2018$year <- c(2018)

#Merge the datasets
fin2014_2015_obama <- rbind(fin2014,fin2015)
fin2017_2018_trump <- rbind(fin2017,fin2018)
```


#### Load pharma stock tickers datasets and merge into one.
https://topforeignstocks.com/stock-lists/the-complete-list-of-biotech-stocks-trading-on-nasdaq/
https://topforeignstocks.com/stock-lists/the-complete-list-of-major-pharmaceutical-stocks-on-the-nyse/

```{r}
stock_tickers <- read.csv("/Users/aselkushkeyeva/Downloads/list_major_pharma.csv")
stock_tickers <- stock_tickers[-48,] 

stock_tickers2 <- read.csv("/Users/aselkushkeyeva/Downloads/list_biotech.csv")
stock_tickers2 <- stock_tickers2[-718,]
stock_tickers2 <- stock_tickers2[, -c(4:14)]

tickers <- rbind(stock_tickers, stock_tickers2)

```


#### make X common and pharmaceuticals in 2014_2015 and 2017_2018 on all sectors data (not subsetted to Healthcare) 
```{r}

# Create a dummy variable to check if X is in tickers dataset created in the previous chunk:
fin2014_2015_obama$X_ph <- ifelse(fin2014_2015_obama$X %in% tickers$Ticker, 1, 0)
fin2017_2018_trump$X_ph <- ifelse(fin2017_2018_trump$X %in% tickers$Ticker, 1, 0)

#Filter pharmaceuticals:
obama_pharma <- fin2014_2015_obama %>% filter(X_ph == 1)
trump_pharma <- fin2017_2018_trump %>% filter(X_ph == 1)

# Find common X for pharmaceuticals because obama_pharma and trump_pharma are of different number of rows:
com <- intersect(obama_pharma$X, trump_pharma$X)
# Create yet another dummy variable to check if X is common:
obama_pharma$X_2 <- ifelse(obama_pharma$X %in% com, 1, 0)
trump_pharma$X_2 <- ifelse(trump_pharma$X %in% com, 1, 0)

#Filter common Xs:
obama_pharma <- obama_pharma %>% filter(X_2 == 1)
trump_pharma <- trump_pharma %>% filter(X_2 == 1)

```

## OBAMA DATA

Some data cleaning:

```{r}

# change all columns class into factor and remove the columns with 1 level factor

obama_ph_fac <- obama_pharma %>% mutate_if(is.numeric,as.factor)
obama_ph_fac<-obama_ph_fac[, sapply(obama_ph_fac, nlevels) > 1]

# change them back into numeric:
obama_ph_num <- obama_ph_fac %>% mutate_if(is.factor,as.numeric)

#impute NAs with mean :

for(i in 1:ncol(obama_ph_num)) {
  obama_ph_num[ , i][is.na(obama_ph_num[ , i])] <- mean(obama_ph_num[ , i], na.rm = TRUE)
}

```
#### Correlation analysis for Obama dataset

```{r}
ob_corrMat <- cor(obama_ph_num, method = "pearson") 

```

```{r}
library(RCurl) # getURL 
library(leaps) # all subsets regression
library(corrplot)
library(caret)
library(FNN)
library(mlbench)
# remove highly correlated
ob_highlyCorr <- findCorrelation(ob_corrMat, cutoff=0.75)
print(ob_highlyCorr)
ob_noncor <- obama_ph_num[,-ob_highlyCorr]  #keep only those not highly correlated
dim(ob_noncor)    ### 113 vars after highly correlated vars are removed + 1 var added below
# correlationMatrix2 <- cor(noncor, method = "pearson")  ### only numeric vars
# correlationMatrix2

ob_noncor$X201X.PRICE.VAR.... <- obama_ph_num$X201X.PRICE.VAR....
dim(ob_noncor) 
```

#### Normalizing the dataset

```{r}
normalize <- function(x) {
               return ((x - min(x)) / (max(x) - min(x))) }
ob_noncor_norm <- as.data.frame(lapply(ob_noncor, normalize))
```



## TRUMP DATA

Some data cleaning:

```{r}

# change all columns class into factor and remove the columns with 1 level factor

trump_ph_fac <- trump_pharma %>% mutate_if(is.numeric,as.factor)
trump_ph_fac<-trump_ph_fac[, sapply(trump_ph_fac, nlevels) > 1]

# change them back into numeric:
trump_ph_num <- trump_ph_fac %>% mutate_if(is.factor,as.numeric)

#impute NAs with mean :

for(i in 1:ncol(trump_ph_num)) {
  trump_ph_num[ , i][is.na(trump_ph_num[ , i])] <- mean(trump_ph_num[ , i], na.rm = TRUE)
}

```
#### Correlation analysis for Trump Dataset

```{r}
tr_corrMat <- cor(trump_ph_num, method = "pearson") 

```

```{r}
library(RCurl)
library(leaps)
library(corrplot)
library(caret)
library(FNN)
library(mlbench)
# remove highly correlated
tr_highlyCorr <- findCorrelation(tr_corrMat, cutoff=0.75)
tr_noncor <- trump_ph_num[,-ob_highlyCorr] 
dim(tr_noncor)    ### 113 vars after highly correlated vars are removed + 1 var added below


tr_noncor$X201X.PRICE.VAR.... <- trump_ph_num$X201X.PRICE.VAR....
dim(tr_noncor) 
```


# Regression Models
#### Forward selection method on Obama data
```{r}
ob_norm <- ob_noncor_norm
ob_norm $X201X.PRICE.VAR.... <- NULL
library(MASS)

full_obama <- lm(Class~., data= ob_norm)
null_obama <- lm(Class~1,data=ob_norm)
stepF_obama <- stepAIC(null_obama, scope=list(lower=null_obama, upper=full_obama), direction= "forward")
summary(stepF_obama)
```


#### Obama Multiple linear regression

```{r}
set.seed(1)
lm_rn_train <- sample(nrow(ob_noncor_norm), floor(nrow(ob_noncor_norm)*0.7))
lm_train <- ob_noncor_norm[lm_rn_train,]
lm_test <- ob_noncor_norm[-lm_rn_train,]
```


```{r}
ob_model_mlr <- lm(X201X.PRICE.VAR....~., data=lm_train) 
#summary(ob_model_mlr)
lm_prediction <- predict(ob_model_mlr, newdata =lm_test)  
```

```{r}
lm_errors <- lm_prediction - lm_test$X201X.PRICE.VAR....
hist(lm_errors)
plot(lm_prediction, lm_test$X201X.PRICE.VAR....)
```

```{r}
lm_rmse <- sqrt(mean((lm_test$X201X.PRICE.VAR.... - lm_prediction)^2))

lm_rel_change <- abs(lm_errors) / lm_test$X201X.PRICE.VAR....
lm_pred25 <- table(lm_rel_change<0.25)["TRUE"] / nrow(lm_test)
lm_pred25
paste("RMSE:", lm_rmse)
paste("PRED(25):", round(lm_pred25,2))

#results:
#     TRUE 
# 0.4618474 
# [1] "RMSE: 0.164229465213062"
# [1] "PRED(25): 0.46"

```

#### Trump Multiple linear regression

```{r}
 #set.seed(1)
#lm_rn_train <- sample(nrow(ob_noncor_norm), floor(nrow(ob_noncor_norm)*0.7))
fin_lm_train <- ob_noncor_norm
fin_lm_test <- tr_noncor_norm
```


```{r}
final_model_mlr <- lm(X201X.PRICE.VAR....~., data=fin_lm_train) 
fin_lm_prediction <- predict(final_model_mlr, newdata =fin_lm_test)  
```

```{r}
fin_lm_errors <- fin_lm_prediction - fin_lm_test$X201X.PRICE.VAR....
#hist(lm_errors)
```

Root mean square error and find the percentage of cases with less than 25% error.
```{r}
fin_lm_rmse <- sqrt(mean((fin_lm_test$X201X.PRICE.VAR.... - fin_lm_prediction)^2))

fin_lm_rel_change <- abs(fin_lm_errors) / fin_lm_test$X201X.PRICE.VAR....
fin_lm_pred25 <- table(fin_lm_rel_change<0.25)["TRUE"] / nrow(fin_lm_test)  ## gives the count of those who are true on the condition of rel_change<0.25
##OR pred25 <- sum((rel_change<0.25)=="TRUE")/nrow(test)
fin_lm_pred25
paste("RMSE:", fin_lm_rmse)
paste("PRED(25):", round(fin_lm_pred25,2))

```

# KNN Models
#### Obama KNN model

```{r}
#divide into train and test
set.seed(1)
ob_rn_train <- sample(nrow(ob_noncor_norm), floor(nrow(ob_noncor_norm)*0.7))  #create train and test data sets
ob_train <- ob_noncor_norm[ob_rn_train,]  
ob_test <- ob_noncor_norm[-ob_rn_train,]
ob_train_labelsKNN <- ob_train[,114]   ##DV in the training set
ob_test_labelsKNN <- ob_test[,114]     ##DV in the test set
ob_train_KNN <- ob_train[,-114]   ##Only keep IV in the training set
ob_test_KNN <- ob_test[,-114]     ##Only keep IV in the test set
```
 
best value of K is 8:
```{r} 
x <- 0
for (i in 1:sqrt(nrow(ob_train)))   #try 1:11 to see the elbow clearly
{
set.seed(1)
ob_KNNmodel <- knn.reg(train =ob_train_KNN , test = ob_test_KNN, y = ob_train_labelsKNN , k = i)  
ob_predicted <- ob_KNNmodel$pred
  
ob_rmse <- sqrt(mean((ob_test_labelsKNN-ob_predicted)^2))
x[i] <- ob_rmse
print(paste(i,ob_rmse))  ##only for the sake of the example
}

plot(x, type="l", col="red")
which.min(x)
```
KNN model:
```{r}
set.seed(1)
ob_KNNmodel <- knn.reg(train =ob_train_KNN , test = ob_test_KNN, y = ob_train_labelsKNN , k = 8)  
ob_predicted <- ob_KNNmodel$pred
plot(ob_test_labelsKNN, ob_predicted, xlab="y", ylab=expression(hat(y)))
ob_errors <- ob_predicted - ob_test_labelsKNN
ob_rmse <- sqrt(mean((ob_test_labelsKNN-ob_predicted)^2))
ob_rmse
ob_rel_change <- abs(ob_errors) / ob_test_labelsKNN
ob_pred25 <- table(ob_rel_change<0.25)["TRUE"] / nrow(ob_test) 
ob_pred25
```

#### Trump KNN model

```{r}
tr_noncor_norm <- as.data.frame(lapply(tr_noncor, normalize))
fin_train_labelsKNN <- ob_noncor_norm[,114]
fin_test_labelsKNN <- tr_noncor_norm[,114]
fin_train <- ob_noncor_norm[,-114]
fin_test <- tr_noncor_norm[,-114]
set.seed(1)
final_KNNmodel <- knn.reg(train =fin_train , test = fin_test, y = fin_train_labelsKNN , k = 8)  
final_predicted <- final_KNNmodel$pred
plot(fin_test_labelsKNN, final_predicted, xlab="y", ylab=expression(hat(y)))
fin_errors <- final_predicted - fin_test_labelsKNN
final_rmse <- sqrt(mean((fin_test_labelsKNN-final_predicted)^2))
final_rmse
fin_rel_change <- abs(fin_errors) / fin_test_labelsKNN
fin_pred25 <- table(fin_rel_change<0.25)["TRUE"] / nrow(fin_test) 
fin_pred25

```


# Classification model for question 2 part 2


# Data Cleaning for Classification
```{r stocktickers found in 200+dataset}
#run this
nyseticker<-read.csv("nyseticker.csv")


nasdaqticker<-read.csv("nasdaqticker.csv")


#change names of pricevar column so it is same for all dataframes
st2002014<-read.csv("2014_Financial_Data.csv")
colnames(st2002014)[224]<-"NxtYearPriceVar"

st2002015<-read.csv("2015_Financial_Data.csv")
colnames(st2002015)[224]<-"NxtYearPriceVar"

st2002016<-read.csv("2016_Financial_Data.csv")
colnames(st2002016)[224]<-"NxtYearPriceVar"

st2002017<-read.csv("2017_Financial_Data.csv")
colnames(st2002017)[224]<-"NxtYearPriceVar"

st2002018<-read.csv("2018_Financial_Data.csv")
colnames(st2002018)[224]<-"NxtYearPriceVar"
unique(st2002014$Sector)


obamastocks<-rbind(st2002014,st2002015)
obamahealthcare<-obamastocks[obamastocks$Sector=="Healthcare", ]

trumpstocks<-rbind(st2002017,st2002018)
trumphealthcare<-trumpstocks[trumpstocks$Sector=="Healthcare", ]

trumphealthcare$nyselabel<-ifelse(trumphealthcare$X%in%nyseticker$Ticker,1,0)
trumphealthcare$naslabel<-ifelse(trumphealthcare$X%in%nasdaqticker$Ticker,1,0)
obamahealthcare$nyselabel<-ifelse(obamahealthcare$X%in%nyseticker$Ticker,1,0)
obamahealthcare$naslabel<-ifelse(obamahealthcare$X%in%nasdaqticker$Ticker,1,0)

```


# Data preperation for Obama dataset

```{r}
#str(obamahealthcare)
#get rid of columns that will give feature selection trouble
obamahealth<-obamahealthcare[,-c(93,112,114,223)]
```

```{r impute missing values}
for(i in 1:ncol(obamahealth)) {
  obamahealth[ , i][is.na(obamahealth[ , i])] <- mean(obamahealth[ , i], na.rm = TRUE)
}
#label the dataset in the data cleaning rmd
#now only pharmaceutical on the stock exchanges are included
notpharma<-c(which(obamahealth$naslabel==0 & obamahealth$nyselabel==0))
obamahealth<-obamahealth[-notpharma,]
sum(is.na(obamahealth))
```
```{r correlation and normalization}
#log everything to normalize dataset, add high constant 
normfunc<-function(x){
  (x-min(x))/(max(x)-min(x))
}
#include range to not get rid of any columns
obamahealth[,2:220]<-lapply(obamahealth[,2:220],normfunc)
numobama<-obamahealth[,2:220]
ObamaCorMat <- cor(numobama, method = "pearson")  ### only numeric vars
highlyCorrelated <- findCorrelation(ObamaCorMat, cutoff=0.75)
obamahealth <- obamahealth[,-highlyCorrelated]
rownames(obamahealth)<-NULL# to reset index

```

```{r backward selection}

#impute missing values to prevent errors in feature selection
#normalize the data
#do correlation analysis first
#make sure to only run models on numeric values
set.seed(1)
full <- lm(Class~.,data=obamahealth)
stepB <- stepAIC(full, direction= "backward", trace=FALSE)
summary(stepB)

```

```{r}
selectedobama<-obamahealth[ ,c(10,14,22,24,27,56,66,85,94,100)]
#rename class labels for proper glm
selectedobama$Class<-ifelse(selectedobama$Class==1,"inc","dec")
selectedobama$Class<-as.factor(selectedobama$Class)

```

```{r}
set.seed(1)
sampletrain <- sample(nrow(selectedobama), floor(nrow(selectedobama)*0.7))
selobtrain <- selectedobama[sampletrain,]#training set with labels
selobtest <- selectedobama[-sampletrain,]#test set with labels
selobtrainlabel<-selobtrain[,10]
selobtestlabel<-selobtest[,10]
selobtest<-selobtest[,-10]#test set no labels
```

# prep trump test data

```{r}
#first run feature selection
trumphealth<-trumphealthcare[,-c(93,112,114,223)]

```

```{r impute missing values}
for(i in 1:ncol(trumphealth)) {
  trumphealth[ , i][is.na(trumphealth[ , i])] <- mean(trumphealth[ , i], na.rm = TRUE)
}
#label the dataset in the data cleaning rmd
#now only pharmaceutical on the stock exchanges are included
notpharma<-c(which(trumphealth$naslabel==0 & trumphealth$nyselabel==0))
trumphealth<-trumphealth[-notpharma,]
sum(is.na(trumphealth))
```

```{r}
#normalize dataset
trumphealth[,2:220]<-lapply(trumphealth[,2:220],normfunc)
#significant features from obama dataset are only included for the trump dataset

selectedtrumpfromob<-trumphealth[ ,c(41,51,72,76,84,155,168,205,215,221)]

selectedtrumpfromob$Class<-ifelse(selectedtrumpfromob$Class==1,"inc","dec")
selectedtrumpfromob$Class<-as.factor(selectedtrumpfromob$Class)
rownames(selectedtrumpfromob)<-NULL #reset index

```

```{r trump features from obama}
set.seed(1)
#sampletrain <- sample(nrow(selectedobama), floor(nrow(selectedobama)*0.7))
fullobtrain <- selectedobama[,]#training set with labels
fulltrumptest <- selectedtrumpfromob[,]#test set with labels
fullobamatrainlabel<-fullobtrain[,10]
fulltrumptestlabel<-fulltrumptest[,10]
fulltrumptest<-fulltrumptest[,-10]#test set no labels
```


# Ensemble Classification model
```{r}
library(caretEnsemble)
set.seed(1)
control <- trainControl(method="repeatedcv", number = 10, repeats=5, savePredictions="final", classProbs = TRUE)
algorithmList <- c('rf', 'svmPoly', 'glm', 'naive_bayes')
set.seed(1)
models_selob <- caretList(Class~., data=selobtrain, trControl=control, methodList=algorithmList)
set.seed(1)
results_selob <- resamples(models_selob)
summary(results_selob)
dotplot(results_selob)  
modelCor(results_selob)

```

```{r}
#no model is very correlated but naive bayes has low accuracy so it was removed
library(caretEnsemble)
set.seed(1)
control2 <- trainControl(method="repeatedcv", number = 10, repeats=5, savePredictions="final",classProbs = TRUE)
algorithmList2 <- c('rf', 'glm', 'svmPoly')
set.seed(1)
models_selob2 <- caretList(Class~., data=selobtrain, trControl=control2, methodList=algorithmList2)
set.seed(1)
results_selob2 <- resamples(models_selob2)
summary(results_selob2)
dotplot(results_selob2)  
modelCor(results_selob2)
```

```{r}
set.seed(1)
stackControl <- trainControl(method="repeatedcv", number=10, repeats=5, savePredictions=TRUE, classProbs=TRUE)
set.seed(1)
selobstack.rf <- caretStack(models_selob2, method="rf", metric="Accuracy", trControl=stackControl)
print(selobstack.rf)  ##ensemble model
```

```{r}
stack.pred <- predict(selobstack.rf , selobtest)
cf_ensemble_selob <- confusionMatrix(as.factor(stack.pred), as.factor(selobtestlabel) , positive="inc", mode = "everything")
print(cf_ensemble_selob)
```

# obama train, trump test 

### Obama model tested with full 2017-2018 data
```{r 70%obama model with fulltrump test}
OTstack.pred <- predict(selobstack.rf , fulltrumptest)
cf_ensemble_OT <- confusionMatrix(as.factor(OTstack.pred), as.factor(fulltrumptestlabel) , positive="inc", mode = "everything")
print(cf_ensemble_OT)
```
